### 2021-04-26 R/DraftFunctions.Rmd
'datacheckformat' function & its test

---
title: "DraftFunctions"
author: "Vincyane Badouard"
date: "26/04/2021"
output: html_document
---

# Fonction de contrôle & formatage de l’INPUT “data” de la fonction pilot (“datacheckformat”):
INPUT: data = Inventaires botaniques Guyafor/ONF (data frame)
OUTPUT: data = Inventaires botaniques type “Guyafor”(data frame)

vérification du type d’inventaire : 
* Test présence des variables.
* Test class des variables.
Si les tests précédents ont tous reçu =TRUE, STOP de la fonction dataformat. 
Sinon mettre en forme “inventaire Guyafor” (nom des var, leurs classes, unités (mettre des limites avec range),...)
re-vérification du type d’inventaire : 
* Test présence des variables
* Test class des variables
Si les tests précédent ont tous reçu =TRUE, STOP de la fonction dataformat.
Sinon Message : “Votre format de données renseignés à l’argument “data”, n’est pas compatible. Voir la vignette du package pour y trouver le format requis.”

```{r}
data("Paracou6_2016") #data in the environment
lapply(Paracou6_2016, class) #give the class of alla the data variables
```

```{r Create a broken dataset}
#in inst/extdata
BrokenParacou6_2016 <- Paracou6_2016 %>% 
  select(-Species) %>% # missing column
  rename(DBH = CircCorr) %>% # other column name
  rename(PLOT = Plot) %>% # variable name in upper case
  mutate(idTree = as.character(as.numeric(idTree))) %>% #an other variable class
  mutate(Lon = round(Lon, digits =0)) %>% # no digits
  mutate(Genus = toupper(Genus)) # values in upper case
```

```{r My tests}

'pouet' %in% names(Paracou6_2016)
if (!is.null(Paracou6_2016$pouet)) 
    stop("pouet parameter should be an integer (column name), or NULL if no presence column.")

```

```{r}
# Variables presence check
"Plot" %in% names(Paracou6_2016)
"CensusYear" %in% names(Paracou6_2016)
"idTree" %in% names(Paracou6_2016)
"Family" %in% names(Paracou6_2016)
"Genus" %in% names(Paracou6_2016)
"Species" %in% names(Paracou6_2016)
"CircCorr" %in% names(Paracou6_2016)
"CodeAlive" %in% names(Paracou6_2016)
"CommercialSp" %in% names(Paracou6_2016)
"UTMZone" %in% names(Paracou6_2016)
'Lat' %in% names(Paracou6_2016)
'Lon' %in% names(Paracou6_2016)
'VernName' %in% names(Paracou6_2016)
'Xfield' %in% names(Paracou6_2016)
'Yfield' %in% names(Paracou6_2016)
'Xutm' %in% names(Paracou6_2016)
'Yutm' %in% names(Paracou6_2016)

```

```{r}
# Variables class check
 if (!inherits(Paracou6_2016$idTree, "integer") && !is.null(idTree)) #majuscule ?
    stop("idTree parameter should be an integer (column name), or NULL if no presence column.")

if (!inherits(Paracou6_2016$Plot, "character") && !is.null(Plot))
    stop("Plot parameter should be a character (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Xfield, "numeric") && !is.null(Xfield))
    stop("Xfield parameter should be numeric (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Yfield, "numeric") && !is.null(Yfield))
    stop("Yfield parameter should be numeric (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Xutm, "numeric") && !is.null(Xutm))
    stop("Xutm parameter should be numeric (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Yutm, "numeric") && !is.null(Yutm))
    stop("Yutm parameter should be numeric (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$UTMZone, "integer") && !is.null(UTMZone))
    stop("UTMZone parameter should be an integer (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Lat, "numeric") && !is.null(Lat))
    stop("Lat parameter should be numeric (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Lon, "numeric") && !is.null(Lon))
    stop("Lon parameter should be numeric (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Family, "character") && !is.null(Family))
    stop("Family parameter should be a character (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Genus, "character") && !is.null(Genus))
    stop("Genus parameter should be a character (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$Species, "character") && !is.null(Species))
    stop("Species parameter should be a character (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$VernName, "character") && !is.null(VernName))
    stop("VernName parameter should be a character (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$CommercialSp, "logical") && !is.null(CommercialSp))
    stop("CommercialSp parameter should be logical (column name), or NULL if no presence column.")

   if (!inherits(Paracou6_2016$CensusYear, "integer") && !is.null(CensusYear))
    stop("CensusYear parameter should be an integer (column name), or NULL if no presence column.")

  if (!inherits(Paracou6_2016$CodeAlive, "logical") && !is.null(CodeAlive))
    stop("CodeAlive parameter should be logical (column name), or NULL if no presence column.")

  if (!inherits(Paracou6_2016$CircCorr, "numeric") && !is.null(CircCorr)) #DBH
    stop("CircCorr parameter should be numeric (column name), or NULL if no presence column.")

# 
# if (!inherits(Paracou6_2016$Forest, "character"))
#     stop("Forest parameter should be a character (column name), or NULL if no presence column.") #pas besoin
#    if (!inherits(Paracou6_2016$idVern, "integer") && !is.null(idVern))
#     stop("idVern parameter should be an integer (column name), or NULL if no presence column.") #pas besoin
# 
#   
#    
#    if (!inherits(Paracou6_2016$PlotArea, "numeric") && !is.null(PlotArea))
#     stop("PlotArea parameter should be numeric (column name), or NULL if no presence column.") #pas besoin
#   
#    if (!inherits(Paracou6_2016$SubPlot, "integer") && !is.null(SubPlot))
#     stop("SubPlot parameter should be an integer (column name), or NULL if no presence column.") #pas besoin
#   
#    if (!inherits(Paracou6_2016$TreeFieldNum, "numeric") && !is.null(TreeFieldNum))
#     stop("TreeFieldNum parameter should be numeric (column name), or NULL if no presence column.") #pas besoin
#   
#    if (!inherits(Paracou6_2016$Project, "character") && !is.null(Project))
#     stop("Project parameter should be a character (column name), or NULL if no presence column.") #pas besoin
#   
#    if (!inherits(Paracou6_2016$Protocole, "character") && !is.null(Protocole))
#     stop("Protocole parameter should be a character (column name), or NULL if no presence column.") #pas besoin
```

```{r all with error message}
if(!all("Plot" %in% names(Paracou6_2016),
"CensusYear" %in% names(Paracou6_2016),
"idTree" %in% names(Paracou6_2016),
"Family" %in% names(Paracou6_2016),
"Genus" %in% names(Paracou6_2016),
"Species" %in% names(Paracou6_2016),
"CircCorr" %in% names(Paracou6_2016),
"CodeAlive" %in% names(Paracou6_2016),
"CommercialSp" %in% names(Paracou6_2016),
"UTMZone" %in% names(Paracou6_2016),
'Lat' %in% names(Paracou6_2016),
'Lon' %in% names(Paracou6_2016),
'VernName' %in% names(Paracou6_2016),
'Xfield' %in% names(Paracou6_2016),
'Yfield' %in% names(Paracou6_2016),
'Xutm' %in% names(Paracou6_2016),
'Yutm' %in% names(Paracou6_2016),
if (!inherits(Paracou6_2016$idTree, "integer") && !is.null(idTree)) #majuscule ?
    stop("idTree parameter should be an integer (column name), or NULL if no presence column."),

if (!inherits(Paracou6_2016$Plot, "character") && !is.null(Plot))
    stop("Plot parameter should be a character (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Xfield, "numeric") && !is.null(Xfield))
    stop("Xfield parameter should be numeric (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Yfield, "numeric") && !is.null(Yfield))
    stop("Yfield parameter should be numeric (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Xutm, "numeric") && !is.null(Xutm))
    stop("Xutm parameter should be numeric (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Yutm, "numeric") && !is.null(Yutm))
    stop("Yutm parameter should be numeric (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$UTMZone, "integer") && !is.null(UTMZone))
    stop("UTMZone parameter should be an integer (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Lat, "numeric") && !is.null(Lat))
    stop("Lat parameter should be numeric (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Lon, "numeric") && !is.null(Lon))
    stop("Lon parameter should be numeric (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Family, "character") && !is.null(Family))
    stop("Family parameter should be a character (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Genus, "character") && !is.null(Genus))
    stop("Genus parameter should be a character (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$Species, "character") && !is.null(Species))
    stop("Species parameter should be a character (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$VernName, "character") && !is.null(VernName))
    stop("VernName parameter should be a character (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$CommercialSp, "logical") && !is.null(CommercialSp))
    stop("CommercialSp parameter should be logical (column name), or NULL if no presence column."),

   if (!inherits(Paracou6_2016$CensusYear, "integer") && !is.null(CensusYear))
    stop("CensusYear parameter should be an integer (column name), or NULL if no presence column."),

  if (!inherits(Paracou6_2016$CodeAlive, "logical") && !is.null(CodeAlive))
    stop("CodeAlive parameter should be logical (column name), or NULL if no presence column."),

  if (!inherits(Paracou6_2016$CircCorr, "numeric") && !is.null(CircCorr)) #DBH
    stop("CircCorr parameter should be numeric (column name), or NULL if no presence column.")))
    stop("Your data format, as specified in the 'data' argument, is not compatible. See the package vignette for the required format.") 
```