### 2021-06-16 R/DESIGNATION_DRAFT.Rmd
.futurereserve

---
title: "DESIGNATION_DRAFT"
author: "Vincyane Badouard"
date: "04/06/2021"
output: html_document
---

# Les arguments
inventory
speciescriteria
type
fuel
objective
diversification
specieslax
objectivelax
otherloggingparameters = loggingparameters()
multilignes "MainTrail"

# Fonctions internes à exporter
harvestable, selected, futurereserve

# Etapes préliminaires
+ Ne sélectionner que des arbres encore vivants (DeathCause = NULL) durant toute la fonction.
+ Identification à vue d'œil des arbres à défaut (proportion : valeur fixe ou modèle ?) : coder "1" dans une colonne "Defect" de la table, coder "0" les autres.
+ Ne considérer dans la désignation que les arbres "0" dans "Defect".
+ Désignation : fonction "TreeSelection"
+ Si type = "RIL3fuelhollow", "VO" = "objective"
+ Sinon "VO" = "objective" * PlotArea + 20-30% pour compenser le bois creux désigné.

```{r Give values to the arguments to work}

inventory = addtreedim(cleaninventory(inventorycheckformat(Paracou6_2016))) #3539 rows (Paracou: 3620 rows)
speciescriteria = SpeciesCriteria
type = "manual"
fuel = "0"
objective = 20
diversification = TRUE
specieslax = FALSE
objectivelax = FALSE
otherloggingparameters = loggingparameters()
# MainTrail

```

```{r}

# inventory <- inventory %>% 
# filter(DeathCause = NULL) %>%  # select only still alived trees (after MainTrails) NULL ou NA

# mutate(VisibleDefectProba = otherloggingparameters$VisiblyRottenModel(DBH)) %>%
# mutate(VisibleDefect = sample(c(0,1), size = 1, replace = F, prob = VisibleDefectProba)) # 1 = default tree, 0 = no visible default (là ma fct ne sait pas qd elle doit mettre 0 ou 1)

# filter(VisibleDefect = "0") %>%

if (type == "RIL3fuelhollow"| (type == "manual"& fuel =="2")){VO = objective #3604 rows
}else{
  VO = (objective + (objective*((otherloggingparameters$ObjectiveBonus)/100))) * unique(inventory$PlotArea) # to compensate for the designated hollow wood.
}

(20 + (20*0.3))*6.25
(20*6.25) + ((20*6.25)*0.3) #same

```

# Joindre les taxonomies ONF/Guyafor (.ONFGuyafortaxojoin)
```{r}

speciescriteria = SpeciesCriteria

genuscriteria <- speciescriteria %>% # for economic names that relate to entire genus
  filter(Species == "spp")

speciescriteria <- speciescriteria %>%
  filter(Species != "spp")

inventoryA <- inventory %>% 
  left_join(genuscriteria, by = "Genus", suffix = c("", ".genus")) %>%
  select(-Species.genus) %>% 
  left_join(speciescriteria, by = c("Genus","Species"), suffix = c(".genus", ".species")) %>% 
  dplyr::rename(VernName = VernName.genus) %>% 
  mutate(ONFName = ifelse(is.na(VernName.species), VernName.genus.genus, VernName.species)) %>% 
  select(-VernName.species, -VernName.genus.genus) %>% 
  mutate(Commercial = ifelse(is.na(Commercial.species), Commercial.genus, Commercial.species)) %>%
  mutate(MinFD = ifelse(is.na(MinFD.species), MinFD.genus, MinFD.species)) %>%
  mutate(UpMinFD = ifelse(is.na(UpMinFD.species), UpMinFD.genus, UpMinFD.species)) %>%
  mutate(MaxFD = ifelse(is.na(MaxFD.species), MaxFD.genus, MaxFD.species)) %>% 
  select(-Commercial.species, -Commercial.genus, -MinFD.species, -MinFD.genus, -UpMinFD.species, -UpMinFD.genus, -MaxFD.species, -MaxFD.genus)

ConcernedRows <- which(inventoryA$ONFName == "maho rouge" & (inventoryA$ScientificName == "Lecythis_poiteaui"| inventoryA$ScientificName == "Lecythis_praeclara" | inventoryA$ScientificName == "Lecythis_holcogyne" | inventoryA$ScientificName == "Lecythis_pneumatophora"| inventoryA$ScientificName == "Lecythis chartacea"| inventoryA$ScientificName == "Lecythis zabucajo"))
if(length(ConcernedRows) > 0){
  inventoryA$ONFName[ConcernedRows] <- NA
  inventoryA$Commercial[ConcernedRows] <- NA
  inventoryA$MinFD[ConcernedRows] <- NA
  inventoryA$UpMinFD[ConcernedRows] <- NA
  inventoryA$MaxFD[ConcernedRows] <- NA
}

ConcernedRows <- which(inventoryA$ONFName == "balata blanc" & (inventoryA$ScientificName == "Micropholis_melinoniana"| inventoryA$ScientificName == "Micropholis_egensis" | inventoryA$ScientificName == "Micropholis_cayennensis" | inventoryA$ScientificName == "Micropholis_obscura"))
if(length(ConcernedRows) > 0){
  inventoryA$ONFName[ConcernedRows] <- NA
  inventoryA$Commercial[ConcernedRows] <- NA
  inventoryA$MinFD[ConcernedRows] <- NA
  inventoryA$UpMinFD[ConcernedRows] <- NA
  inventoryA$MaxFD[ConcernedRows] <- NA
}

ConcernedRows <- which(inventoryA$ONFName == "kimboto" & (inventoryA$ScientificName == "Pradosia_cochlearia"| inventoryA$ScientificName == "Pradosia_huberi"))
if(length(ConcernedRows) > 0){
  inventoryA$ONFName[ConcernedRows] <- NA
  inventoryA$Commercial[ConcernedRows] <- NA
  inventoryA$MinFD[ConcernedRows] <- NA
  inventoryA$UpMinFD[ConcernedRows] <- NA
  inventoryA$MaxFD[ConcernedRows] <- NA
}

filter(inventoryA, ScientificName == "Micropholis_obscura")

# Special cases:
# Maho rouge = tous les Lecythis (sauf Lecythis poiteaui, Lecythis praeclara et Lecythis holcogyne , Lecythis pneumatophora et Lecythis chartacea, Lecythis zabucajo) 
# Balata blanc = Tous les Micropholis (sauf Micropholis melinoniana, Micropholis egensis,  Micropholis cayennensis et M. obscura)
# Kimboto = tous les Pradosia sauf Pradosia cochlearia et P. huberi

# cèdre blanc = Ocotea guianensis, O. canaliculata, O. oblonga et O. argyrophylla.
# cèdre noir = Ocotea tomentella, Ocotea schomburgkiana, O. glomerata et O. ceanothifolia.
# Dodomissinga = Parkia decussata, P. gigantocarpa, P. nitida, P. reticulata, P. ulei et P. velutina.
# Cèdre gris = les Lauraceae qui ne sont pas d'autres cèdres ...
```

```{r}
# To control
dupli <- inventoryA[duplicated(inventoryA$idTree), ] # les lignes dupliquées sont dues aux genres qui concernent plsrs essences

NAtable <- inventoryA[is.na(inventoryA$Commercial),] #2101 ind not linked
NAtablescfc <- as.data.frame(unique(inventoryA[is.na(inventoryA$Commercial),]$ScientificName)) #264 sp not linked
NAtablevern <- as.data.frame(unique(inventoryA[is.na(inventoryA$Commercial),]$VernName)) #122 sp not linked

NAtablevern <- unique(NAtable$VernName) #264 
NAtablescfc <- unique(NAtable$ScientificName) #122

filter(inventoryA, Commercial =="1") #177 ind
filter(inventoryA, Commercial =="2") #1267 ind
comerc <- filter(inventoryA, Commercial =="1"| Commercial =="2")
all(comerc$CommercialSp == TRUE)
```

# harvestable
+ Étiqueter "harvestable", dans une colonne "LoggingStatus", sélection des individus :
Essences :
- Si diversification=T ou si diversification=F & specieslax=T : = "1" & "2" dans colonne "Commercial"
- Sinon (diversification=F & specieslax=F): = "1" dans colonne "Commercial"
Diamètre exploitable (speciescriteria data): individus dont le DBH est compris entre la valeur dans colonne "MinFD" pour leur sp, de la table speciescriteria , et le MaxFD.
Distribution spatiale:
- Arbres sur <22% de pente et arbres sur pente >22% si à 40 m maximum d’une zone <22%.
- arbre non isolé : élimination des arbres à >100m des autres individus de la même espèce.
- hors des pistes principales : dont les coordonnées n'appartiennent pas aux multilignes "MainTrail".
+ Si diversification=F & specieslax=T, étiqueter "harvestable2nd" à la place d’"harvestable", les "Commercial"= "2" dans la colonne "LoggingStatus".
+ HVinit= somme des valeurs de "TreeHarvestableVolume " des arbres dont "LoggingStatus" = "harvestable".

```{r}
# Il manque les scénarios dans les conditions -_-

#select essences
HarverstableConditions <- # = 1 boolean vector
  if (diversification || (!diversification && specieslax)) {
    inventoryA$Commercial =="1"| inventoryA$Commercial == "2" # now or maybe after we will diversify
  } else if (!diversification && !specieslax) {
    Commercial == "1" # We will never diversify
  }

#select diameters
HarverstableConditions <- HarverstableConditions & (inventoryA$DBH >= inventoryA$MinFD & inventoryA$DBH <= inventoryA$MaxFD) # harvestable species, accord by their DBH

#select spatially
# HarverstableConditions <- HarverstableConditions & ()
##slope
##isolement
##MainTrails out

inventoryB <- inventoryA %>% 
  mutate(Commercial = ifelse(is.na(Commercial), "0", Commercial)) %>% #To avoid NA problems
  mutate(LoggingStatus = ifelse(HarverstableConditions, #Under the above criteria, designate the harvestable species
                                "harvestable", "non-harvestable")) %>% 
  mutate(LoggingStatus = ifelse(Commercial == "0", #The non-commercial species are non-harvestable.
                                "non-harvestable", LoggingStatus)) %>%

  mutate(LoggingStatus = ifelse( 
    !diversification & 
      specieslax & #designate the secondarily harvestable species, because diversification only if necessary
      LoggingStatus == "harvestable" & 
      Commercial == "2",
    "harvestable2nd", LoggingStatus))

HarvestableTable <- inventoryB %>%
  filter(LoggingStatus == "harvestable") #96 ind
HVinit <- sum(HarvestableTable$TreeHarvestableVolume) #compute the harvestable volume in the plot for these criteria
#all the inventory: HVinit = 2088.052 m3
#just harvestable: HVinit = 365.3774 m3

```

# selected
```{r}
# if objective achieved at the first attempt
inventoryC <- inventoryB %>% 
  mutate(Selected = ifelse(HVinit == VO & LoggingStatus == "harvestable",
                           "1", NA))# if we have our volume, harvestable sp = selected sp

inventoryC <- add_column(inventoryC, Up = 0) # Create a column to indicate Which sp is FD uped. Or create this column d'office in speciescriteria. "0" = no uped, "1"= uped. dans speciescriteria ou inventory ?

if (HVinit < VO){ #diversification is necessary, designate the secondary-economic-rank species too
  if (!diversification && specieslax){  
    inventoryD <- inventoryC %>%
      mutate(Condition = ifelse(LoggingStatus == "harvestable2nd" | LoggingStatus == "harvestable", TRUE, FALSE)) %>% 
      group_by(Condition) %>% 
      arrange(desc(TreeHarvestableVolume)) %>% 
      mutate(VolumeCumSum = cumsum(TreeHarvestableVolume)) %>% 
      ungroup() %>% 
      mutate(Selected = ifelse(Condition & VolumeCumSum <= VO, "1", NA)) %>%
      select(-Condition)

    HarvestableTable <- inventoryD %>% 
      filter(Selected == "1")
    HVlax <- sum(HarvestableTable$TreeHarvestableVolume) #49.69643 Harvestable volume, with specieslax permission

    if (HVlax < VO && objectivelax) message("The exploitable volume (= ",paste(HVlax),") is still lower (by ",paste(VO- HVlax),") than your objective volume despite the diversification you have allowed (without the diversification HVinit= ",paste(HVinit),"). In this case, you have chosen to continue harvesting with a volume lower than your objective.")

    if (HVlax < VO && !objectivelax) stop("The harvestable volume = ",paste(HVlax),") is still lower (by ",paste(VO- HVlax),") than your objective volume despite the diversification you have allowed (without the diversification HVinit= ",paste(HVinit),"). By default or by your choice, the simulation stops. If you wish to continue the exploitation in spite of an exploitable volume lower than your objective volume, you have the argument 'objectivelax'.")

    if (!HVlax == VO) message("The harvestable volume (=",paste(HVinit),") was lower (by ",paste(VO- HVinit),") than your objective volume. You have chosen to diversify your species selection in this case. The exploitation was therefore carried out on this diversified selection of species.")
  }

  if (!diversification && !specieslax && objectivelax)
    message("The harvestable volume (=",paste(HVinit),") is less (by ",paste(VO- HVinit),") than your objective volume. In this case you have chosen to continue logging without diversifying your species.")

  if (diversification && objectivelax)
    message("The harvestable volume (=",paste(HVinit),") is less (by ",paste(VO- HVinit),") than your objective volume. In this case you have chosen to continue logging.")

  if ((!specieslax & !objectivelax) | (diversification && !objectivelax)) 
    stop("The harvestable volume (=",paste(HVinit),") is lower (by ",paste(VO- HVinit),") than your objective volume. By default or by your choice, the simulation stops. If you wish to continue the exploitation in spite of a harvestable volume lower than your objective volume, you can use the argument 'objectivelax' or the diversification of species (if it is not already the case).")
}

if (HVinit > VO) {

  inventoryD <- inventoryC %>%
    mutate(LoggingStatus = ifelse(
      Commercial == "1" & (DBH >= UpMinFD & DBH <= MaxFD), #designate preferred individuals of first economic rank species, when the plot is species-rich.
      "harvestableUp", LoggingStatus)) %>%

    mutate(Up = ifelse(Commercial == "1", "1", Up))# to inform that the "1" ranks have been FD upgraded. Pas mieux de le faire à l'sp?
  # test "Commercial"= "1" pour les "harvestableUp"

  if (!diversification) {
    HarvestableTable <- inventoryD %>%
      filter(LoggingStatus == "harvestableUp")

    HVupCommercial1 <- sum(HarvestableTable$TreeHarvestableVolume) #82.42823. compute the harvestable volume with upgraded FD individuals

    if (HVupCommercial1 == VO){

      inventoryE <- inventoryD %>%
        mutate(Selected = ifelse(HVupCommercial1 == VO & LoggingStatus == "harvestableUp", "1", NA))# if harvestableUp individuals are sufficient to have our volume, harvestableUp ind = selected ind

      message("As the harvestable volume (=",paste(HVinit),") was higher (by ",paste(VO- HVinit),") than the objective volume, the Minimum Falling Diameter (MinFD) of 1st economic rank species were increased. The objective volume has now been reached.")
    }

    if (HVupCommercial1 > VO){
      # only individuals with DBH > FD are taken, but not all because their volumes > VO
      inventoryE<- inventoryD %>%
        mutate(Condition = ifelse(LoggingStatus == "harvestableUp", TRUE, FALSE)) %>%
        group_by(Condition) %>%
        arrange(desc(TreeHarvestableVolume)) %>% 
        mutate(VolumeCumSum = cumsum(TreeHarvestableVolume)) %>%
        ungroup() %>%
        mutate(Selected = ifelse(Condition & VolumeCumSum <= VO, "1", NA)) %>% 
        select(-Condition)

      HarvestableTable <- inventoryE %>%
        filter(Selected == "1")
      HVupCommercial1adjust <- sum(HarvestableTable$TreeHarvestableVolume) #49.69643 Harvestable volume, with "1" rank species and upgraded FD individuals only

      message("The harvestable volume (=",paste(HVupCommercial1),") is always higher (by ",paste(VO- HVupCommercial1),") than your objective volume despite the increase in Minimum Falling Diameter (MinFD) (Initial harvestable volume = HVinit). In order to reach your objective volume, the trees were selected in decreasing order of volume until the objective volume was reached.")
    }
    if (HVupCommercial1 < VO){

      inventoryE<- inventoryD %>%
        mutate(Condition = ifelse(LoggingStatus == "harvestableUp" | LoggingStatus == "harvestable", TRUE, FALSE)) %>%
        group_by(Condition) %>%
        arrange(desc(TreeHarvestableVolume)) %>% 
        mutate(VolumeCumSum = cumsum(TreeHarvestableVolume)) %>%
        ungroup() %>%
        mutate(Selected = ifelse(Condition & VolumeCumSum <= VO, "1", NA)) %>% 
        select(-Condition)

      HarvestableTable <- inventoryE %>%
        filter(Selected == "1")
      HVupCommercial1adjust <- sum(HarvestableTable$TreeHarvestableVolume) #49.69643 Harvestable volume, with "1" rank species and upgraded FD individuals

      message("As the harvestable volume (=",paste(HVinit),") was higher (by ",paste(VO- HVinit),") than the objective volume, the Minimum Falling Diameter (MinFD) of the 1st economic rank species were increased with retention of some stems with DBH lower than the UpMinFD to ensure that the objective volume was attained.")
    }
  }

  if (diversification) {

    HarvestableTable <- inventoryD %>%
      filter(LoggingStatus == "harvestableUp" | LoggingStatus == "harvestable") #the upgraded 1st rank and normal 2nd rank.

    HVupCommercial1 <- sum(HarvestableTable$TreeHarvestableVolume) #173.334 m3

    if (HVupCommercial1 == VO){
      inventoryE <- inventoryD %>%
        mutate(Selected = ifelse(LoggingStatus == "harvestableUp" | LoggingStatus == "harvestable",
                                 "1", NA))# if we have our volume, harvestable ind = selected ind
      message("As the harvestable volume (=",paste(HVinit),") was higher (by ",paste(VO- HVinit),") than the objective volume, the Minimum Falling Diameter (MinFD) of the 1st economic rank species were increased. The objective volume has now been reached. It was not necessary to increase the MinFD of the other economic species")
    }

    if (HVupCommercial1 < VO){

      inventoryE<- inventoryD %>%
        mutate(Condition = ifelse(LoggingStatus == "harvestableUp" | LoggingStatus == "harvestable", TRUE, FALSE)) %>%
        group_by(Condition) %>%
        arrange(desc(TreeHarvestableVolume)) %>% 
        mutate(VolumeCumSum = cumsum(TreeHarvestableVolume)) %>%
        ungroup() %>%
        mutate(Selected = ifelse(Condition & VolumeCumSum <= VO, "1", NA)) %>% 
        select(-Condition)

      HarvestableTable <- inventoryE %>%
        filter(Selected == "1")
      HVupCommercial1adjust <- sum(HarvestableTable$TreeHarvestableVolume) #49.69643 Harvestable volume, with "1" rank species and upgraded FD individuals, and 2nd rank no upgraded FD individuals

      message("As the harvestable volume (=",paste(HVinit),") was higher (by ",paste(VO- HVinit),") than the objective volume, the Minimum Falling Diameter (MinFD) of the 1st economic rank species were increased with retention of some stems with DBH lower than the UpMinFD to ensure that the objective volume was attained. It was not necessary to raise the MinFDs of other economic species.")
    }

    if (HVupCommercial1 > VO){

      inventoryE <- inventoryD %>%
        mutate(LoggingStatus = ifelse(LoggingStatus == "harvestable" & Commercial == "2" & DBH >= UpMinFD, #designate preferred individuals of 2nd economic rank species too, when the plot is species-rich.
      "harvestableUp", LoggingStatus)) %>%

        mutate(Up = ifelse(Commercial == "2", "1", Up))# to inform that the "2" ranks have been FD upgraded. Pas mieux de le faire à l'sp?
      HarvestableTable <- inventoryE %>%
        filter(LoggingStatus == "harvestableUp")

      HVupCommercial12 <- sum(HarvestableTable$TreeHarvestableVolume) #271.2342 Harvestable volume, with upgraded FD individuals

      if (HVupCommercial12 == VO){
        inventoryF <- inventoryE %>%
          mutate(Selected = ifelse(LoggingStatus == "harvestableUp",
                                   "1", NA))# if we have our volume, harvestable ind = selected ind
      }
      if (HVupCommercial12 != VO){

        inventoryF <- inventoryE %>%
          mutate(Condition = ifelse(LoggingStatus == "harvestableUp" | LoggingStatus == "harvestable", TRUE, FALSE)) %>%
          group_by(Condition) %>%
          arrange(desc(TreeHarvestableVolume)) %>%
          mutate(VolumeCumSum = cumsum(TreeHarvestableVolume)) %>%
          ungroup() %>%
          mutate(Selected = ifelse(Condition & VolumeCumSum <= VO, "1", NA)) %>%
          select(-Condition)

        HarvestableTable <- inventoryF %>%
          filter(Selected == "1")
        HVupCommercial12adjust <- sum(HarvestableTable$TreeHarvestableVolume) #45.44885 Harvestable volume, with upgraded FD individuals
      }

      message("As the harvestable volume (=",paste(HVinit),") was higher (by ",paste(VO- HVinit),") than the objective volume, it was necessary to increase the Minimum Falling Diameter (MinFD) of all species. The objective volume has now been reached.")

    }
  }
}

# Apply S. Schmitt's "Rotten" predictive model to identify "truly" hollow designated trees.
inventoryG <- inventoryF %>% 
  mutate(ProbedHollowProba = ifelse(Selected == "1", otherloggingparameters$RottenModel(DBH), NULL)) %>%  #Estimates the probability of being probed hollow

  # generate either "1" or "0" randomly for each line, depending on the proba associated with the line:
  rowwise() %>%
  mutate(ProbedHollow = ifelse(!is.na(ProbedHollowProba), sample(c(1,0), size = 1, replace = F, prob = c(ProbedHollowProba, 1-ProbedHollowProba)), NA)) %>%  # 1 = hollow tree, 0 = not hollow
  mutate(ProbedHollow = factor(as.numeric(ProbedHollow))) %>% 
  mutate(Selected = ifelse(ProbedHollow == "1", "0", Selected)) %>%  #hollow probed trees are deselected
   #non-upgraded MinFD species:
  mutate(Up = ifelse(is.na(Up) , "0", Up))

if (type != "RIL3fuelhollow"| (type == "manual"& fuel !="2"))
  HarvestableTable <- inventoryG %>%
  filter(Selected == "1")
VolumewithHollowslost <- sum(HarvestableTable$TreeHarvestableVolume) #128.5047. Harvestable volume, with Hollows lost
VO - VolumewithHollowslost #34 m3 lost: the bonus is therefore generous here (37.5 m3 of bonus).

if (type == "RIL3fuelhollow"| (type == "manual"& fuel =="2")) {
  # Hollow trees = fuelwood:
  inventoryH <- inventoryG %>%
    mutate(DeathCause = ifelse(ProbedHollow == "1", "hollowfuel", NA)) # remplacer NA par DeathCause dans le simulateur ONF

  # Create 2 points vectors with coordinates of the probed hollow trees: "HollowTreesPoints" and "EnergywoodTreesPoints":
HollowTreescoord <- inventoryG %>% 
  filter(ProbedHollow == "1") %>% 
  select(Xutm, Yutm)

HollowTreesPoints  <- st_multipoint(x = as.matrix(HollowTreescoord))
EnergywoodTreesPoints <- HollowTreesPoints

}

```

# futurereserve

```{r}
#Future: select essence and diameters

inventoryH <- inventoryG %>%
  mutate(LoggingStatus = ifelse(Commercial == "1" &
                                  (DBH >= otherloggingparameters$FutureTreesMinDiameter & DBH < MinFD),
                                "future", LoggingStatus))

#Reserve
# Randomly select the reserved trees (among the futures, as many as the number of trees exploited):

StemNbr <- sum(as.numeric(inventoryH$Selected == "1"), na.rm = TRUE)#29 selected ind 
ReserveRows <- sample(which(inventoryH$LoggingStatus == "future"), size = StemNbr, replace = F)

inventoryH$LoggingStatus[ReserveRows] <-"reserve"

```

# Fin treeselection function
```{r}
# Créer un vecteur ("HarvestableTreesPoints") de points de coordonnées des arbres LoggingStatus = "harvestable" ou "harvestableUp" ou "harvestable2nd".
# Créer un vecteur ("SelectedTreesPoints") de points de coordonnées des arbres Selected = 1.
# Créer un vecteur ("FutureTreesPoints") de points de coordonnées des arbres LoggingStatus = "future"
# Créer un vecteur ("ReserveTreesPoints") de points de coordonnées des arbres LoggingStatus = "reserve".

#where specieslax was not necessary, consider eco2s as non-exploitable:
inventoryI <- inventoryH %>% 
  mutate(LoggingStatus = ifelse(HVinit > VO & LoggingStatus == "harvestable2nd", "non-harvestable", LoggingStatus))

unique(inventoryI$LoggingStatus)

write.csv2(inventoryI, "D:/VSC Kourou/DATA/PostDesignationInventory.csv")

```