### 2021-07-20 R/testaddtreedim_DRAFT.Rmd

---
title: "testaddtreedim_DRAFT"
author: "Vincyane Badouard"
date: "04/06/2021"
output: html_document
---

```{r}
if (!("DBH" %in% names(Paracou6_2016))) {add_column(Paracou6_2016, DBH = NA) #if DBH doesn't exist create it
Paracou6_2016$DBH = Paracou6_2016$CircCorr/pi} # and compute it

inventory = Paracou6_2016
crowndiameterparameters = ParamCrownDiameterAllometry
volumeparameters = ForestZoneVolumeParametersTable
advancedloggingparameters = loggingparameters()

spParamCrownDiameter <- crowndiameterparameters %>% #parameters at species scale
  filter(Taxo == "sp") %>%
  select(-Family)
genParamCrownDiameter <- crowndiameterparameters %>% #parameters at genus scale
  filter(Taxo == "gen") %>%
  select(-Family, -ScientificName, -Species)
famParamCrownDiameter <- crowndiameterparameters %>% #parameters at family scale
  filter(Taxo == "fam") %>%
  select(-Genus, -ScientificName, -Species)

# Variables computation:

prout <- inventory %>%

  # TreeHarvestableVolume (m3)
  left_join(volumeparameters) %>%
  mutate(TreeHarvestableVolume = advancedloggingparameters$TreeHarvestableVolumeAllometry(aCoef, bCoef, DBH)) %>%

  # TrunkHeight (m)
  mutate(TrunkHeight = advancedloggingparameters$TrunkHeightAllometry(DBH, TreeHarvestableVolume)) %>%

  # TreeHeight (m)
  mutate(TreeHeight = advancedloggingparameters$TreeHeightAllometry(DBH)) %>%

  # CrownHeight (m)
  mutate(CrownHeight = TreeHeight - TrunkHeight) %>%

# CrownDiameter (m)
# Add crown diameter allometry parameters at the inventory:
unite(Genus, Species, col = "ScientificName", sep = "_", remove = F) %>% # create ScientificName column in the inventory
  left_join(spParamCrownDiameter, by = c("ScientificName","Genus","Species")) %>% # at species scale
  left_join(genParamCrownDiameter, by = "Genus", suffix = c(".species", ".genus")) %>% # at genus scale
  left_join(famParamCrownDiameter, by = "Family") %>% # at family scale
  dplyr::rename(Taxo.family = Taxo, alpha.family = alpha, beta.family = beta) %>%
  # manage the different scales parameters Species > Genus > Family:
  # alpha:
  mutate(alpha = alpha.species, beta = beta.species, Taxo = Taxo.species) %>% # create the future & unique alpha,beta,taxo columns
  mutate(alpha = ifelse(is.na(alpha), alpha.genus, alpha)) %>% #if species parameters are absent, take genus parameters
  mutate(alpha = ifelse(is.na(alpha), alpha.family, alpha)) %>% #if species&genus parameters are absent, take family parameters
  mutate(alpha = ifelse(is.na(alpha), mean(alpha, na.rm =T), alpha)) %>% #if species,genus&family parameters are absent, take parameters mean
  select(-alpha.species, -alpha.genus, -alpha.family) %>% #remove obsolete columns
  # beta:
  mutate(beta = ifelse(is.na(beta), beta.genus, beta)) %>% #if species parameters are absent, take genus parameters
  mutate(beta = ifelse(is.na(beta), beta.family, beta)) %>% #if species&genus parameters are absent, take family parameters
  mutate(beta = ifelse(is.na(beta), mean(beta, na.rm =T), beta)) %>% #if species,genus&family parameters are absent, take parameters mean
  select(-beta.species, -beta.genus, -beta.family) %>% #remove obsolete columns
  # scale:
  mutate(Taxo = ifelse(is.na(Taxo), Taxo.genus, Taxo)) %>% #if species parameters are absent, take genus parameters
  mutate(Taxo = ifelse(is.na(Taxo), Taxo.family, Taxo)) %>% #if species&genus parameters are absent, take family parameters
  mutate(Taxo = ifelse(is.na(Taxo), "mean", Taxo)) %>% #if species,genus&family parameters are absent, take parameters mean
  select(-Taxo.species, -Taxo.genus, -Taxo.family) %>% #remove obsolete columns
  # compute the crown diameter
  mutate(CrownDiameter = advancedloggingparameters$CrownDiameterAllometry(DBH, TreeHeight, alpha, beta))

return(inventory)

```