### 2021-08-16 R/Crowns_draft.Rmd
createcanopy and treefromthesky functions #38 #39

---
title: "Crowns_draft"
author: "Vincyane Badouard"
date: "16/08/2021"
output: html_document
---

```{r}
inventory <- addtreedim(inventorycheckformat(Paracou6_2016))
dat <- inventory[679,]
```

```{r}
# The crown
treefromthesky <- function(dat){

  Crown <- dat %>%
    mutate(xCrown = Xutm, # X centroid
           yCrown = Yutm, # Y ventroid
           exCrown = CrownDiameter/2,
           eyCrown = CrownHeight/2) %>%
    st_as_sf(coords = c("xCrown", "yCrown")) # ellipse centroid coordinates
  Crown <- st_ellipse(Crown, Crown$exCrown, Crown$eyCrown) # create the ellipse

}
```

```{r}
library(ggplot2)
g <- ggplot() +
  geom_sf(data = st_as_sf(inventory, coords = c("Xutm", "Yutm")), aes(label = ScientificName)) +
  geom_sf(data = Crown, fill = "forestgreen") # trees polygons

plotly::ggplotly(g)
```

```{r}
Canopy <- inventory %>%
  group_by(idTree) %>% # for each tree
  do(Crowns = # inform geometry. # Filling a column from a function whose input is a table
       treefromthesky(.) %>%
       st_as_text()) %>% # as text to easy join with a non spacial table
  tidyr::unnest(Crowns) # here to pass from list to character

inventory <- left_join(inventory, Canopy, by = "idTree") # join spatial filtered inventory and non spatial complete inventory
```

```{r}
#The small ones first so that they are behind the big ones on the plot
inventory <- arrange(inventory, TreeHeight)

ggplot() +
  geom_sf(data = getgeometry(inventory, Crowns),
          aes(alpha = TreeHeight), # label = paste(idTree, Species), 
          fill = "forestgreen")
```

# Plotly version
```{r}
#The small ones first so that they are behind the big ones on the plot
inventory <- arrange(inventory, TreeHeight) %>% 
  filter(TreeHeight >30)

g <- ggplot() +
  geom_sf(data = getgeometry(inventory, Crowns),
          aes(label = paste(idTree, Species), alpha = 1/2), 
          fill = "forestgreen") # trees polygons

plotly::ggplotly(g)
```