### 2021-08-17 R/DRAFT_secondtrailsopening.Rmd
Maria secondary track WIP 1/2

---
title: "DRAFT_secondtrailopening"
output: html_document
---

```{r include=FALSE}
library(Maria)
library(readr)
library(dplyr)
library(tibble)
library(tidyr)
library(ggplot2)
library(reshape)
library(sf)
library(raster)
```

```{r include=FALSE}
data(DemParacou)
data(Paracou6_2016)
data(ExploitPolygones)

scenario = "RIL1"
advancedloggingparameters = loggingparameters()

DemParacou <- raster(paste0(Sys.getenv('DATAPATH'),"/GIS/Paracou/Topo_P6_PARACOU.tif"))

Plots <- rgdal::readOGR(paste0(Sys.getenv('DATAPATH'),"/GIS/Paracou/Plot_P6_PARACOU.shp"))

MainTrails <- st_as_sf(as(rgdal::readOGR(paste0(Sys.getenv('DATAPATH'),"/GIS/Paracou/maintrail_P6_PARACOU.shp")),"SpatialLines" ))
PlotSlope <- raster::terrain(raster::mask( x = DemParacou,mask = Plots))

scenarios <- scenariosparameters(scenario = scenario, objective = 70,
fuel = "2", diversification = TRUE, winching = NULL, directionalfelling = NULL)

```

```{r}

inventory <- ONFGuyafortaxojoin(addtreedim(cleaninventory(
inventorycheckformat(Paracou6_2016))))

treeselectionoutputs <- treeselection(inventory, objective = scenarios$objective,
scenario =scenario,fuel = scenarios$fuel , diversification = scenarios$diversification, winching = scenarios$winching , specieslax = FALSE,
objectivelax = FALSE, DEM = DemParacou, plotslope = PlotSlope,
exploitpolygones = ExploitPolygones,
speciescriteria = SpeciesCriteria,
advancedloggingparameters = loggingparameters())

inventory <- treeselectionoutputs$inventory

```

```{r}
AccessPolygones <- FilterAccesExplArea(exploitpolygones = ExploitPolygones,maintrails = MainTrails,winching = scenarios$winching,advancedloggingparameters = loggingparameters())

DfDEM <- as_tibble(rasterToPoints(DemParacou)) %>% mutate( z = Topo_P6_PARACOU) %>% dplyr::select(-Topo_P6_PARACOU)

ggplot() + geom_raster(data = DfDEM, aes(x =x , y = y, fill = z)) + 
  geom_sf(data = MainTrails) + geom_sf(data = AccessPolygones,fill = "red",alpha = 0.1) + geom_sf(data = treeselectionoutputs$HarvestableTreesPoints, aes(color = "Harvestable")) + 
  geom_sf(data =treeselectionoutputs$SelectedTreesPoints , aes(color = "Selected")) + geom_sf(data = treeselectionoutputs$ReserveTreesPoints, aes(color = "Reserves")) + 
  geom_sf(data = treeselectionoutputs$FutureTreesPoints, aes(color = "Futures")) +
  geom_sf(data = treeselectionoutputs$BigTreesPoints, aes(color = "DBH>50cm"))+ scale_fill_gradientn(colours = rev(terrain.colors(225)))

```

```{r}

# secondtrailsopening <- function(inventory,
#                                 DEM = DemParacou,
#                                 plotslope =  PlotSlope,
#                                 plots = Plots,
#                                 prospectionunit = ProspectionUnit,
#                                 maintrails = MainTrails,
#                                 selectedtreespoints = SelectedTreesPoints,
#                                 hollowtreespoints = HollowTreesPoints,
#                                 avenirtreespoints = AvenirTreesPoint,
#                                 reservetreespoints = ReserveTreesPoints,
#                                 deadtreespoints = DeadTreesPoints,
#                                 otherlogginparameters = logginparameters()) {

  # Arguments check

  if(!inherits(inventory, "data.frame"))
    stop("The 'inventory' arguments of the 'secondtrailsopening' function must be data.frame")

  if(!inherits(Plots, "sf"))
    stop("The 'plots' arguments of the 'secondtrailsopening' function must be sf polygon")

  if(!any(unlist(lapply(list(DemParacou, PlotSlope), inherits, "RasterLayer"))))
    stop("The 'DEM' and 'plotslope' arguments of the 'secondtrailsopening' function must be raster")

  # Redefinition of the parameters according to the chosen scenario
  scenariosparameters <- scenariosparameters(objective = objective, type = type, fuel = fuel,
                                             diversification = diversification)

  objective <- scenariosparameters$objective
  fuel <- scenariosparameters$fuel
  diversification <- scenariosparameters$diversification

##################################

  CostSlopeMatrix <- list(list(Slope = 3, Cost = 1),
                          list(Slope = 5, Cost = 3),
                          list(Slope = 12, Cost = 20),
                          list(Slope = 22, Cost = 60),
                          list(Slope = 27, Cost = 600),
                          list(Slope = Inf, Cost = 1000))

  AccessMainTrails <- FilterAccesExplArea(exploitpolygones = ExploitPolygones,maintrails = MainTrails,winching = "0",advancedloggingparameters = loggingparameters()) %>%
  sf::st_cast("POLYGON") %>% sf::st_as_sf() %>% 
    mutate(ID = paste0("ID_",row_number()))

  PartMainTrails <- st_intersection(MainTrails  ,
                                    AccessMainTrails %>% 
                                      sf::st_buffer(dist = advancedloggingparameters$MaxMainTrailWidth)) %>% 
    sf::st_buffer(dist = advancedloggingparameters$MaxMainTrailWidth) %>% sf::st_union() %>% 
    sf::st_intersection(sf::st_as_sf(Plots) %>% sf::st_union()) %>%
     sf::st_cast("POLYGON")  %>% 
    sf::st_as_sf() %>% 
    sf::st_join(AccessMainTrails)

  AccessPoint <- PartMainTrails  %>%  
    sf::st_sample( rep(1,dim(PartMainTrails)[1]) ,type = "random",by_polygon = TRUE) %>% sf::st_as_sf() %>%  
    sf::st_join(PartMainTrails)

    CostRaster <- raster(extent(DemParacou),resolution = res(DemParacou), crs = crs(DemParacou))
  values(CostRaster) <- 1000
  CostRaster <- mask(CostRaster, Plots)
  CostRaster <- crop(CostRaster, Plots)

  # RastersToPoints

  plotSlopePoint <- dplyr::as_tibble(raster::rasterToPoints(PlotSlope))

  CostRasterPoint <- dplyr::as_tibble(raster::rasterToPoints(CostRaster))

  # left join par x et y
  plotTib <-
    dplyr::left_join(plotSlopePoint, CostRasterPoint, by = c('x', 'y'))

  CostSlopeRaster <- plotTib %>%
    mutate(Exploit = if_else(
        slope <= atan(CostSlopeMatrix[[1]]$Slope/100), # on est en radians pour la pente, donc .264 radians = 27% de pente
      true = CostSlopeMatrix[[1]]$Cost,
      false = if_else(
        slope <= atan(CostSlopeMatrix[[2]]$Slope/100),
        true = CostSlopeMatrix[[2]]$Cost,
        false = if_else(
          slope <= atan(CostSlopeMatrix[[3]]$Slope/100),
          true = CostSlopeMatrix[[3]]$Cost,
          false = if_else(
            slope <= atan(CostSlopeMatrix[[4]]$Slope/100),
            true = CostSlopeMatrix[[4]]$Cost,
            false = if_else(
              slope <= atan(CostSlopeMatrix[[5]]$Slope/100),
              true = CostSlopeMatrix[[5]]$Cost,
              false = CostSlopeMatrix[[6]]$Cost)
          )
        )
      )
    )) %>% dplyr::select(x,y,Exploit)

  CostRaster <- raster::rasterFromXYZ(CostSlopeRaster, crs = 32622)

    AccessRaster <- raster(extent(DemParacou),resolution = res(DemParacou), crs = crs(DemParacou))
  values(AccessRaster) <- 1000

  AccessRaster <- rasterize(x = as_Spatial(AccessPolygones %>% 
                                             sf::st_buffer(dist =  -advancedloggingparameters$ScndTrailWidth/2)),
                            y = AccessRaster , 
                            field = 0, 
                            update = TRUE) 
    AccessRaster <- crop(AccessRaster,  CostRaster)
  AccessRaster <- mask(AccessRaster, Plots)

  CostRaster<- CostRaster + AccessRaster

  BigTreesRaster <- raster(extent(DemParacou),resolution = res(DemParacou), crs = crs(DemParacou))
  values(BigTreesRaster) <- 0
  BigTreesRaster <- rasterize(x = as_Spatial(treeselectionoutputs$BigTreesPoints %>% 
                                                     sf::st_buffer(dist =  advancedloggingparameters$ScndTrailWidth/2 + 2) ),
                                    y = BigTreesRaster , 
                                    field = 500, 
                                    update = TRUE) 
  BigTreesRaster <- crop(BigTreesRaster,  CostRaster)

    ReserveRaster <- raster(extent(DemParacou),resolution = res(DemParacou), crs = crs(DemParacou))
  values(ReserveRaster) <- 0
  ReserveRaster <- rasterize(x = as_Spatial(treeselectionoutputs$HarvestableTreesPoints %>% 
                                                     sf::st_buffer(dist =  advancedloggingparameters$ScndTrailWidth/2 + 2) ),
                                    y = ReserveRaster , 
                                    field = 500, 
                                    update = TRUE) 
  ReserveRaster <- crop(ReserveRaster,  CostRaster)

  FutureRaster <- raster(extent(DemParacou),resolution = res(DemParacou), crs = crs(DemParacou))
  values(FutureRaster ) <- 0
  FutureRaster  <- rasterize(x = as_Spatial(treeselectionoutputs$HarvestableTreesPoints %>% 
                                                     sf::st_buffer(dist =  advancedloggingparameters$ScndTrailWidth/2 +2) ),
                                    y = FutureRaster  , 
                                    field = 50, 
                                    update = TRUE) 
  FutureRaster  <- crop(FutureRaster ,  CostRaster)

  CostRaster<- CostRaster + BigTreesRaster + ReserveRaster + FutureRaster

  CostRaster <- rasterize(x = as_Spatial(PartMainTrails),
                                    y = CostRaster , 
                                    field = 1E-3, 
                                    update = TRUE)

  CostRasterMean <- aggregate(CostRaster, fact=3, fun=mean)
  CostRasterMean <- crop(CostRasterMean,  CostRaster)
  CostRasterMean <- mask(CostRasterMean, Plots)

  DEM <- crop(DemParacou,  CostRasterMean)
  DEM <- mask(DEM, Plots)

  DEMmean <- aggregate(DEM, fact=3, fun=mean)
  DEMmean <- crop(DEMmean,  CostRasterMean)
  DEMmean <- mask(DEMmean, Plots)

  CondSurf <- 1/CostRasterMean

  SlopeCond <- SlopeRdCond(DEM = DEMmean,fact = 1,advancedloggingparameters = loggingparameters())

  pts <- treeselectionoutputs$SelectedTreesPoints %>% sf::st_cast("POINT") %>% sf::st_join(AccessMainTrails)
  pts <- rbind(AccessPoint %>% mutate(type = "Access"),pts %>% mutate(type = "Tree"))

  pathLines <- list()
  k <- 0

##################################

  if (winching == "0") {

    for (accessid in unique(pts$ID)) {

      selectedPoints <- pts %>% filter(ID == accessid) 
      PointAcc <- selectedPoints %>% filter(type == "Access") %>% 
        mutate(ptAcc = dplyr::row_number()) %>% 
        mutate(EstCost = NA)
      CostDistEst <- AdjTopoLCP(costSurface = CondSurf,
                                DEM = DEMmean , 
                                pts = selectedPoints %>% 
                                  as_Spatial(),
                                slopeRdCond = SlopeCond,
                                paths = FALSE) [,1:dim(PointAcc)[1]]
      CostDistEst[CostDistEst == 0] <- NA
      CostDistEst <- na.exclude(CostDistEst)
      PointTree <- selectedPoints %>% filter(type == "Tree") %>% 
        mutate(ptAcc = max.col(-CostDistEst))
      if (dim(PointAcc)[1] == 1) {
        PointTree <- PointTree %>% mutate(EstCost = CostDistEst)
      }else{PointTree <- PointTree %>% mutate(EstCost = apply(CostDistEst,1, min))}
      selectedPoints <- rbind(PointAcc,PointTree)

      for (accessPtId in 1:dim(PointAcc)[1]) {
        TmpSelectedPts <- selectedPoints %>% 
          filter(ptAcc == accessPtId)

        TmpAccPts <- TmpSelectedPts %>% 
          filter(type == "Access")

        TmpTreePts <- TmpSelectedPts %>% 
          filter(type == "Tree") %>% 
          arrange(desc(EstCost))

        if (dim(TmpTreePts)[1] != 0) {
          for (TreeId in 1:dim(TmpTreePts)[1]){

            TmpPtsWIP <- rbind(TmpAccPts,TmpTreePts[TreeId,])
            TmpPathWIP <-  AdjTopoLCP(costSurface = CondSurf,DEM = DEMmean , pts = TmpPtsWIP %>% as_Spatial(),
                              slopeRdCond = SlopeCond,paths = TRUE)

            CostRasterMean  <- rasterize(x = TmpPathWIP[[2]] ,
                                    y = CostRasterMean  , 
                                    field = 0.3,update =  TRUE)
            CondSurf <- 1/CostRasterMean

            k <- k +1

            pathLines[[k]] <- TmpPathWIP[[2]]
            pathLines[[k]]@lines[[1]]@ID <- paste("Path",accessid,"-",accessPtId, "-", TreeId, sep = " ")

          }
        }
      }

    }

  }

  if (winching == "1") {

    ptsCabble <- pts %>% filter(type == "Tree") %>%
      st_buffer(dist = advancedloggingparameters$CableLength) %>%
      st_intersection() %>% 
      st_make_valid() %>%
      arrange(desc(n.overlaps)) %>% 
      mutate(ID_pts = paste0("T.",row_number()))

    for (accessid in unique(pts$ID)) {

      PointAcc <- pts %>% filter(ID == accessid) %>% filter(type == "Access")%>% 
       mutate(ptAcc = row_number()) %>% 
        mutate(EstCost = NA) %>% 
        mutate(ID_pts = paste0("A.",row_number()))

      InPlot <- st_intersects(ptsCabble %>% 
        filter(ID == accessid) %>% 
        st_centroid(),Plots %>% st_as_sf(),sparse = FALSE)

      ptsCabbleCentroid <- ptsCabble %>% 
        filter(ID == accessid) %>% 
        st_centroid() %>% dplyr::mutate(InPlot = InPlot) %>% 
        filter(InPlot == TRUE) %>% dplyr::select(-InPlot) %>% 
       mutate(ptAcc = NA) %>% 
      mutate(EstCost = NA) %>% arrange(desc(n.overlaps))

      selectedPoints <- rbind(PointAcc %>% mutate(n.overlaps = NA, origins = NA),ptsCabbleCentroid %>% 
        st_sf())

      CostDistEst <- AdjTopoLCP(costSurface = CondSurf,
                                DEM = DEMmean , 
                                pts = selectedPoints %>% 
                                  as_Spatial(),
                                slopeRdCond = SlopeCond,
                                paths = FALSE) [,1:dim(PointAcc)[1]]
      CostDistEst[CostDistEst == 0] <- NA
      CostDistEst <- na.exclude(CostDistEst)
      PointTree <- selectedPoints %>% filter(type == "Tree") %>% 
        mutate(ptAcc = max.col(-CostDistEst))
      if (dim(PointAcc)[1] == 1) {
        PointTree <- PointTree %>% mutate(EstCost = CostDistEst)
      }else{PointTree <- PointTree %>% mutate(EstCost = apply(CostDistEst,1, min))}
      selectedPoints <- rbind(PointAcc %>% mutate(n.overlaps = NA, origins = NA),PointTree)

      for (accessPtId in 1:dim(PointAcc)[1]) {
        TmpSelectedPts <- selectedPoints %>% 
          filter(ptAcc == accessPtId)

        TmpAccPts <- TmpSelectedPts %>% 
          filter(type == "Access")

        TmpTreePts <- TmpSelectedPts %>% 
          filter(type == "Tree") %>% 
          arrange(desc(n.overlaps),EstCost)

        TreeId <- 1

          while (dim(TmpTreePts)[1] != 0){

            TmpPtsWIP <- rbind(TmpAccPts %>%  
                                st_union() %>% 
                                st_cast("POINT")%>% 
                                 st_as_sf() %>% 
                                 mutate(ID = accessid)%>% 
                                 mutate(type = "Access") %>% 
                                 mutate(ptAcc = accessPtId) %>% 
                                 mutate(ID_Pts = NA),ptsCabble %>% 
                                filter(ID_pts == TmpTreePts[TreeId,"ID_pts"]$ID_pts)  %>%  
                                st_union() %>% 
                                st_cast("POINT")%>% 
                                 st_as_sf()%>% 
                                 mutate(ID = accessid)%>% 
                                 mutate(type = "Overlay")%>% 
                                 mutate(ptAcc = accessPtId) %>% 
                                 mutate(ID_Pts = TmpTreePts[TreeId,"ID_pts"]$ID_pts))

            CostDistEstWIP <-  AdjTopoLCP(costSurface = CondSurf,DEM = DEMmean , pts = TmpPtsWIP %>% as_Spatial(),
                              slopeRdCond = SlopeCond,paths = FALSE)[,1]

            CostDistEstWIP[CostDistEstWIP == 0] <- NA
            CostDistEstWIP <- na.exclude(CostDistEstWIP)

            PointTreeWIP <- TmpPtsWIP %>% filter(type == "Overlay") %>% 
              mutate(EstCost= CostDistEstWIP) %>% arrange(EstCost)

            TmpPtsWIP <- rbind(TmpPtsWIP %>% filter(type == "Access") %>% mutate(EstCost = NA),PointTreeWIP[1,])

            TmpPathWIP <- AdjTopoLCP(costSurface = CondSurf,DEM = DEMmean , pts = TmpPtsWIP %>% as_Spatial(),
                              slopeRdCond = SlopeCond,paths = TRUE)

            CostRasterMean  <- rasterize(x = TmpPathWIP[[2]] ,
                                    y = CostRasterMean  , 
                                    field = 0.3,update =  TRUE)
            k <- k +1

            pathLines[[k]] <- TmpPathWIP[[2]]
            pathLines[[k]]@lines[[1]]@ID <- paste("Path",accessid,"-",accessPtId, "-", TreeId, sep = " ")

            CbblePtsWIP <- ptsCabble %>% filter(ID_pts == TmpTreePts[TreeId,"ID_pts"]$ID_pts)

            CbblePtsWIP$origins [[1]] %in% TmpTreePts$origins [[2]]

            TreeId <- TreeId + 1

          }
        }
      }

    }

  }

  if (winching == "2") {

  }

  paths <- do.call(rbind, pathLines)

    df <-  as_tibble(rasterToPoints(CostRasterMean)) %>% mutate( Cost = layer) %>% dplyr::select(-layer)
  df2 <- as_tibble(rasterToPoints(DEM)) %>% mutate( z = Topo_P6_PARACOU) %>% dplyr::select(-Topo_P6_PARACOU)
  df3 <- as_tibble(rasterToPoints(raster(SlopeCond))) %>% mutate( AccSlp = layer) %>% dplyr::select(-layer)
  ggplot() +
    geom_raster(data = df, aes(x =x , y = y, fill = Cost)) +  geom_sf(data = pts) + scale_fill_gradientn(colours = rev(terrain.colors(225)))  + geom_sf( data = sf::st_as_sf(paths))    + geom_sf(data =  paths %>% st_as_sf()  %>% smoothr::smooth(method = "ksmooth",smoothness = 5) %>% st_buffer(dist = 4) %>% st_union(), color = 'red', alpha = 0.1 )  + theme_bw()

```