### 2021-08-23 R/DRAFT_secondtrailsopening (conflicted copy 2021-08-23 005423).Rmd

---
title: "DRAFT_secondtrailopening"
output: html_document
---

```{r include=FALSE}
library(Maria)
library(readr)
library(dplyr)
library(tibble)
library(tidyr)
library(ggplot2)
library(reshape)
library(sf)
library(raster)
```

```{r include=FALSE}
data(DemParacou)
data(Paracou6_2016)
data(ExploitPolygones)

scenario = "RIL3"
advancedloggingparameters = loggingparameters()

DemParacou <- raster(paste0(Sys.getenv('DATAPATH'),"/GIS/Paracou/Topo_P6_PARACOU.tif"))

Plots <- rgdal::readOGR(paste0(Sys.getenv('DATAPATH'),"/GIS/Paracou/Plot_P6_PARACOU.shp"))

MainTrails <- st_as_sf(as(rgdal::readOGR(paste0(Sys.getenv('DATAPATH'),"/GIS/Paracou/maintrail_P6_PARACOU.shp")),"SpatialLines" ))
PlotSlope <- raster::terrain(raster::mask( x = DemParacou,mask = Plots))

scenarios <- scenariosparameters(scenario = scenario, objective = 140,
fuel = "2", diversification = TRUE, winching = NULL, directionalfelling = NULL)

```

```{r}

inventory <- ONFGuyafortaxojoin(addtreedim(cleaninventory(
inventorycheckformat(Paracou6_2016))))

treeselectionoutputs <- treeselection(inventory, objective = scenarios$objective,
scenario =scenario,fuel = scenarios$fuel , diversification = scenarios$diversification, winching = scenarios$winching , specieslax = FALSE,
objectivelax = FALSE, DEM = DemParacou, plotslope = PlotSlope,
exploitpolygones = ExploitPolygones,
speciescriteria = SpeciesCriteria,
advancedloggingparameters = loggingparameters())

inventory <- treeselectionoutputs$inventory

```

```{r}
AccessPolygones <- FilterAccesExplArea(exploitpolygones = ExploitPolygones,maintrails = MainTrails,winching = scenarios$winching,advancedloggingparameters = loggingparameters())

DfDEM <- as_tibble(rasterToPoints(DemParacou)) %>% mutate( z = Topo_P6_PARACOU) %>% dplyr::select(-Topo_P6_PARACOU)

ggplot() + geom_raster(data = DfDEM, aes(x =x , y = y, fill = z)) + 
  geom_sf(data = MainTrails) + geom_sf(data = AccessPolygones,fill = "red",alpha = 0.1) + geom_sf(data = treeselectionoutputs$HarvestableTreesPoints, aes(color = "Harvestable")) + 
  geom_sf(data =treeselectionoutputs$SelectedTreesPoints , aes(color = "Selected")) + geom_sf(data = treeselectionoutputs$ReserveTreesPoints, aes(color = "Reserves")) + 
  geom_sf(data = treeselectionoutputs$FutureTreesPoints, aes(color = "Futures")) +
  geom_sf(data = treeselectionoutputs$BigTreesPoints, aes(color = "DBH>50cm"))+ scale_fill_gradientn(colours = rev(terrain.colors(225)))

```

```{r}

# secondtrailsopening <- function(inventory,
#                                 DEM = DemParacou,
#                                 plotslope =  PlotSlope,
#                                 plots = Plots,
#                                 prospectionunit = ProspectionUnit,
#                                 maintrails = MainTrails,
#                                 selectedtreespoints = SelectedTreesPoints,
#                                 hollowtreespoints = HollowTreesPoints,
#                                 avenirtreespoints = AvenirTreesPoint,
#                                 reservetreespoints = ReserveTreesPoints,
#                                 deadtreespoints = DeadTreesPoints,
#                                 otherlogginparameters = logginparameters()) {

  # Arguments check

  if(!inherits(inventory, "data.frame"))
    stop("The 'inventory' arguments of the 'secondtrailsopening' function must be data.frame")

  if(!inherits(Plots, "sf"))
    stop("The 'plots' arguments of the 'secondtrailsopening' function must be sf polygon")

  if(!any(unlist(lapply(list(DemParacou, PlotSlope), inherits, "RasterLayer"))))
    stop("The 'DEM' and 'plotslope' arguments of the 'secondtrailsopening' function must be raster")

  # Redefinition of the parameters according to the chosen scenario
  scenariosparameters <- scenariosparameters(objective = objective, type = type, fuel = fuel,
                                             diversification = diversification)

  objective <- scenariosparameters$objective
  fuel <- scenariosparameters$fuel
  diversification <- scenariosparameters$diversification

##################################

  CostSlopeMatrix <- list(list(Slope = 3, Cost = 3),
                          list(Slope = 5, Cost = 5),
                          list(Slope = 12, Cost = 20),
                          list(Slope = 22, Cost = 60),
                          list(Slope = 27, Cost = 600),
                          list(Slope = Inf, Cost = 1000))

  AccessMainTrails <- FilterAccesExplArea(exploitpolygones = ExploitPolygones,maintrails = MainTrails,winching = "0",advancedloggingparameters = loggingparameters()) %>%
  sf::st_cast("POLYGON") %>% sf::st_as_sf() %>% 
    mutate(ID = paste0("ID_",row_number()))

  PartMainTrails <- st_intersection(MainTrails  ,
                                    AccessMainTrails %>% 
                                      sf::st_buffer(dist = advancedloggingparameters$MaxMainTrailWidth)) %>% 
    sf::st_buffer(dist = advancedloggingparameters$MaxMainTrailWidth) %>% sf::st_union() %>% 
    sf::st_intersection(sf::st_as_sf(Plots) %>% sf::st_union()) %>%
     sf::st_cast("POLYGON")  %>% 
    sf::st_as_sf() %>% 
    sf::st_join(AccessMainTrails)

  AccessPoint <- PartMainTrails  %>%  
    sf::st_sample( rep(1,dim(PartMainTrails)[1]) ,type = "random",by_polygon = TRUE) %>% sf::st_as_sf() %>%  
    sf::st_join(PartMainTrails)

    CostRaster <- raster(extent(DemParacou),resolution = res(DemParacou), crs = crs(DemParacou))
  values(CostRaster) <- 1000
  CostRaster <- mask(CostRaster, Plots)
  CostRaster <- crop(CostRaster, Plots)

  # RastersToPoints

  plotSlopePoint <- dplyr::as_tibble(raster::rasterToPoints(PlotSlope))

  CostRasterPoint <- dplyr::as_tibble(raster::rasterToPoints(CostRaster))

  # left join par x et y
  plotTib <-
    dplyr::left_join(plotSlopePoint, CostRasterPoint, by = c('x', 'y'))

  CostSlopeRaster <- plotTib %>%
    mutate(Exploit = if_else(
        slope <= atan(CostSlopeMatrix[[1]]$Slope/100), # on est en radians pour la pente, donc .264 radians = 27% de pente
      true = CostSlopeMatrix[[1]]$Cost,
      false = if_else(
        slope <= atan(CostSlopeMatrix[[2]]$Slope/100),
        true = CostSlopeMatrix[[2]]$Cost,
        false = if_else(
          slope <= atan(CostSlopeMatrix[[3]]$Slope/100),
          true = CostSlopeMatrix[[3]]$Cost,
          false = if_else(
            slope <= atan(CostSlopeMatrix[[4]]$Slope/100),
            true = CostSlopeMatrix[[4]]$Cost,
            false = if_else(
              slope <= atan(CostSlopeMatrix[[5]]$Slope/100),
              true = CostSlopeMatrix[[5]]$Cost,
              false = CostSlopeMatrix[[6]]$Cost)
          )
        )
      )
    )) %>% dplyr::select(x,y,Exploit)

    CostRaster <- raster::rasterFromXYZ(CostSlopeRaster, crs = 32622)

    AccessRaster <- raster(extent(DemParacou),resolution = res(DemParacou), crs = crs(DemParacou))
  values(AccessRaster) <- 1000

  AccessRaster <- rasterize(x = as_Spatial(AccessPolygones %>% 
                                             sf::st_buffer(dist =  -advancedloggingparameters$ScndTrailWidth/2)),
                            y = AccessRaster , 
                            field = 0, 
                            update = TRUE) 
    AccessRaster <- crop(AccessRaster,  CostRaster)
  AccessRaster <- mask(AccessRaster, Plots)

  CostRaster<- CostRaster + AccessRaster

  BigTreesRaster <- raster(extent(DemParacou),resolution = res(DemParacou), crs = crs(DemParacou))
  values(BigTreesRaster) <- 0
  BigTreesRaster <- rasterize(x = as_Spatial(treeselectionoutputs$BigTreesPoints %>% 
                                                     sf::st_buffer(dist =  advancedloggingparameters$ScndTrailWidth/2 + 2) ),
                                    y = BigTreesRaster , 
                                    field = 500, 
                                    update = TRUE) 
  BigTreesRaster <- crop(BigTreesRaster,  CostRaster)

    ReserveRaster <- raster(extent(DemParacou),resolution = res(DemParacou), crs = crs(DemParacou))
  values(ReserveRaster) <- 0
  ReserveRaster <- rasterize(x = as_Spatial(treeselectionoutputs$HarvestableTreesPoints %>% 
                                                     sf::st_buffer(dist =  advancedloggingparameters$ScndTrailWidth/2 + 2) ),
                                    y = ReserveRaster , 
                                    field = 500, 
                                    update = TRUE) 
  ReserveRaster <- crop(ReserveRaster,  CostRaster)

  FutureRaster <- raster(extent(DemParacou),resolution = res(DemParacou), crs = crs(DemParacou))
  values(FutureRaster ) <- 0
  FutureRaster  <- rasterize(x = as_Spatial(treeselectionoutputs$HarvestableTreesPoints %>% 
                                                     sf::st_buffer(dist =  advancedloggingparameters$ScndTrailWidth/2 +2) ),
                                    y = FutureRaster  , 
                                    field = 50, 
                                    update = TRUE) 
  FutureRaster  <- crop(FutureRaster ,  CostRaster)

  CostRaster<- CostRaster + BigTreesRaster + ReserveRaster + FutureRaster

    if (scenarios$winching == "2") {

  plotSlopePointGrpl <- dplyr::as_tibble(raster::rasterToPoints(PlotSlope))

  CostRasterPointGrpl <- dplyr::as_tibble(raster::rasterToPoints(CostRaster))

  # left join par x et y
  plotTibGrpl <-
    dplyr::left_join(plotSlopePointGrpl, CostRasterPointGrpl, by = c('x', 'y'))

  CostSlopeRasterGrpl <- plotTibGrpl %>%
    mutate(Exploit = if_else(
        slope <= atan(advancedloggingparameters$GrappleMaxslope/100), # on est en radians pour la pente, donc .264 radians = 27% de pente
      true = 0,
      false = Inf
    )) %>% dplyr::select(x,y,Exploit)

  CostRasterGrpl <- raster::rasterFromXYZ(CostSlopeRasterGrpl, crs = 32622)

  CostRasterGrpl <- CostRaster + CostRasterGrpl

  CostRasterGrpl <- rasterize(x = as_Spatial(PartMainTrails),
                                    y = CostRasterGrpl , 
                                    field = 1E-4, 
                                    update = TRUE)

  CostRasterMeanGrpl <- aggregate(CostRasterGrpl, fact=3, fun=mean)
  CostRasterMeanGrpl <- crop(CostRasterMeanGrpl,  CostRaster)
  CostRasterMeanGrpl <- mask(CostRasterMeanGrpl, Plots)

  }

    CostRaster <- rasterize(x = as_Spatial(PartMainTrails),
                                    y = CostRaster , 
                                    field = 1E-4, 
                                    update = TRUE)

  CostRasterMean <- aggregate(CostRaster, fact=3, fun=mean)
  CostRasterMean <- crop(CostRasterMean,  CostRaster)
  CostRasterMean <- mask(CostRasterMean, Plots)

  DEM <- crop(DemParacou,  CostRasterMean)
  DEM <- mask(DEM, Plots)

  DEMmean <- aggregate(DEM, fact=3, fun=mean)
  DEMmean <- crop(DEMmean,  CostRasterMean)
  DEMmean <- mask(DEMmean, Plots)

  CondSurf <- 1/CostRasterMean

  SlopeCond <- SlopeRdCond(DEM = DEMmean,fact = 1,advancedloggingparameters = loggingparameters())

  pathLines <- list()
  k <- 1

  pts <- treeselectionoutputs$SelectedTreesPoints %>% 
    sf::st_cast("POINT") %>% 
    mutate(ID = NA) %>% 
    mutate(type = "Tree")

##################################

  winching <- scenarios$winching

  if (winching == "0") {

  CostDistEst <- AdjTopoLCP(costSurface = CondSurf,
                                DEM = DEMmean , 
                                pts = pts %>% 
                                  as_Spatial(),
                                slopeRdCond = SlopeCond,
                                paths = FALSE) [,1:dim(AccessPoint)[1]]
  CostDistEst[CostDistEst == 0] <- NA
  CostDistEst <- na.exclude(CostDistEst)
  PointTree <- pts %>% filter(type == "Tree") %>% 
      mutate(ptAcc = max.col(-CostDistEst))
  PointTree$ID <- as.factor(PointTree$ptAcc)
  levels(PointTree$ID) <- as.character(AccessPoint$ID)

  pts <- rbind(AccessPoint %>% 
                 mutate(type = "Access") %>% 
                 mutate(ptAcc = row_number()) %>% 
                 mutate(IPpts = paste0("A.",row_number())),
               PointTree %>% 
                 mutate(IPpts = paste0("T.",row_number())))

  PointAcc <- pts %>% filter(type == "Access") %>% 
        mutate(EstCost = NA)

  CostDistEst <- AdjTopoLCP(costSurface = CondSurf,
                                DEM = DEMmean , 
                                pts = pts %>% 
                                  as_Spatial(),
                                slopeRdCond = SlopeCond,
                                paths = FALSE) [,1:dim(PointAcc)[1]]
  CostDistEst[CostDistEst == 0] <- NA
  CostDistEst <- na.exclude(CostDistEst)
  PointTree <- pts %>% filter(type == "Tree") %>% 
        mutate(ptAcc = max.col(-CostDistEst))
  if (dim(PointAcc)[1] == 1) {
      PointTree <- PointTree %>% mutate(EstCost = CostDistEst)
    }else{PointTree <- PointTree %>% mutate(EstCost = apply(CostDistEst,1, min))}
  selectedPoints <- rbind(PointAcc,PointTree)

  for (accessPtId in 1:dim(PointAcc)[1]) {
    TmpSelectedPts <- selectedPoints %>% 
      filter(ptAcc == accessPtId)

    TmpAccPts <- TmpSelectedPts %>% 
      filter(type == "Access")

    TmpTreePts <- TmpSelectedPts %>% 
      filter(type == "Tree") %>% 
      arrange(desc(EstCost))

    if (dim(TmpTreePts)[1] != 0) {
      for (TreeId in 1:dim(TmpTreePts)[1]){

        TmpPtsWIP <- rbind(TmpAccPts,TmpTreePts[TreeId,])
        TmpPathWIP <-  AdjTopoLCP(costSurface = CondSurf,DEM = DEMmean , pts = TmpPtsWIP %>% as_Spatial(),
                              slopeRdCond = SlopeCond,paths = TRUE)

        CostRasterMean  <- rasterize(x = TmpPathWIP[[2]] ,
                                    y = CostRasterMean  , 
                                    field = 0.1,update =  TRUE)
        CondSurf <- 1/CostRasterMean

        pathLines[[k]] <- TmpPathWIP[[2]]
        pathLines[[k]]@lines[[1]]@ID <- paste("Path",accessPtId, "-", TreeId, sep = " ")

        k <- k +1

      }
    }
   }
  }else{

    # winching == 1/2

    PointAcc <- AccessPoint %>% #def Access Point
      mutate(type = "Access") %>% 
      mutate(IDpts = paste0("A.",row_number())) %>% 
      mutate(n.overlaps = NA, origins = NA) %>% 
      dplyr::select(-ID)

    if (winching == "2") {
      ptsGrpl <- pts %>% filter(type == "Tree") %>% #def Grpl point
                  st_buffer(dist = advancedloggingparameters$GrappleLength) %>%
                  st_set_precision(1) %>% 
                  st_intersection() %>% 
                  st_make_valid()

       ptsGrpl <- ptsGrpl %>%  filter(st_intersects(ptsGrpl ,AccessMainTrails %>% st_union(),sparse = FALSE)) %>% 
                  mutate(IDpts = paste0("I.",row_number()))

       ptsWIP <- ptsGrpl %>% #def Grpl point as WIP point
                  st_centroid() %>% 
                  dplyr::select(-ID)

       ptsCbl <- pts %>% filter(type == "Tree") %>% #def cbl point
                  st_buffer(dist = advancedloggingparameters$CableLength) %>%
                  st_set_precision(1) %>% 
                  st_intersection() %>% 
                  st_make_valid()

        ptsCbl <- ptsCbl %>% 
                  filter(st_intersects(ptsCbl ,AccessMainTrails %>% st_union() ,sparse = FALSE)) %>% 
                  mutate(IDpts = paste0("I.",row_number()))

        ptsWIPCbl <-  ptsCbl %>%
          st_centroid() %>%
          dplyr::select(-ID)

        ptsWIP <- ptsWIP %>% 
                  arrange(desc(n.overlaps))

        RemainTree <- dim(ptsWIP %>% filter(type == "Tree"))[1] + dim(ptsWIPCbl %>% filter(type == "Tree"))[1]

    }else{
        ptsCbl <- pts %>% filter(type == "Tree") %>% #def cbl point
                  st_buffer(dist = advancedloggingparameters$CableLength) %>%
                  st_set_precision(1) %>% 
                  st_intersection() %>% 
                  st_make_valid() %>% 
                  mutate(IDpts = paste0("I.",row_number()))

        ptsWIP <-  ptsCbl %>%#def cbl point as WIP point
          st_centroid() %>%
          dplyr::select(-ID)

        ptsWIP <- ptsWIP %>%  #filter cbl intersection centroid point out plot
                  filter(st_intersects(ptsWIP ,Plots %>% st_as_sf(),sparse = FALSE)) %>% 
                  arrange(desc(n.overlaps))

        RemainTree <- dim(ptsWIP %>% filter(type == "Tree"))[1]

    }

    limRec <- dim(pts)[1]

    WinchingSwitch <- FALSE

    while (RemainTree !=0L & k <= limRec) {

      Grpl2CblFlag <- FALSE

      print(paste0("iter : ",k, "| Max : ",limRec) )

    ptsWIPmax <- rbind(PointAcc,ptsWIP %>% 
                         filter(n.overlaps == max(ptsWIP$n.overlaps))) %>% 
              mutate(TypeAcc = NA) %>% 
              mutate(EstCost = NA)  %>% 
              mutate(ptsAcc = NA)

    CostDistEstCbl <- AdjTopoLCP(costSurface = CondSurf,
                                DEM = DEMmean , 
                                pts = ptsWIPmax %>% 
                                  as_Spatial(),
                                slopeRdCond = SlopeCond,
                                paths = FALSE) [,1:dim(PointAcc)[1]]

    CostDistEstCbl <- CostDistEstCbl[(dim(PointAcc)[1]+1):dim(CostDistEstCbl)[1],1:dim(PointAcc)[1]]

    PointTree <- ptsWIPmax %>% filter(type == "Tree") %>% 
        mutate(ptAccCbl = max.col(-matrix(CostDistEstCbl,ncol = dim(PointAcc)[1] ),ties.method = "first"))

    if (dim(PointAcc)[1] == 1) {
      PointTree <- PointTree %>% mutate(EstCostCbl = matrix(CostDistEstCbl,ncol = dim(PointAcc)[1] ))
    }else{
      PointTree <- PointTree %>% mutate(EstCostCbl = apply(CostDistEstCbl,1, min))
      }

      if (winching == "2") {

        CondSurfGrpl <- 1/CostRasterMeanGrpl

        CostDistEstGrpl <- AdjTopoLCP(costSurface = CondSurfGrpl,
                                DEM = DEMmean , 
                                pts = ptsWIPmax %>% 
                                  as_Spatial(),
                                slopeRdCond = SlopeCond,
                                paths = FALSE) [,1:dim(PointAcc)[1]]

        CostDistEstGrpl <- CostDistEstGrpl[(dim(PointAcc)[1]+1):dim(CostDistEstGrpl)[1],1:dim(PointAcc)[1]]

        PointTree <- PointTree %>% 
          mutate(ptAccGpl = max.col(-t(CostDistEstGrpl),ties.method = "first"))

          if (dim(PointAcc)[1] == 1) {
            PointTree <- PointTree %>% mutate(EstCostGrpl = CostDistEstGrpl)
          }else{
              PointTree <- PointTree %>% mutate(EstCostGrpl = apply(CostDistEstGrpl,1, min))}

        CostDistEstGrpl[CostDistEstGrpl != Inf] <- 0
        CostDistEstGrpl[CostDistEstGrpl == Inf] <- 1

        for (j in 1:dim(CostDistEstGrpl)[1]) {
          PointTree[j,"TypeAcc"] <-  as.character(prod(CostDistEstGrpl[j,]))
        }
        PointTree$TypeAcc[PointTree$TypeAcc == "0"] <- "Grpl"
        PointTree$TypeAcc[PointTree$TypeAcc == "1"] <- "Cbl"

        PointTree$ptsAcc[PointTree$TypeAcc == "Grpl"] <- PointTree$ptAccGpl[PointTree$TypeAcc == "Grpl"]
        PointTree$ptsAcc[PointTree$TypeAcc != "Grpl"] <- PointTree$ptAccCbl[PointTree$TypeAcc != "Grpl"]

        PointTree$EstCost[PointTree$TypeAcc == "Grpl"] <- PointTree$EstCostGrpl[PointTree$TypeAcc == "Grpl"]
        PointTree$EstCost[PointTree$TypeAcc != "Grpl"] <- PointTree$EstCostCbl[PointTree$TypeAcc != "Grpl"]

        PointTree <- PointTree %>% arrange(desc(TypeAcc),EstCost)

            if (PointTree$TypeAcc[1] == "Grpl") {

              TmpPtsWIP <- ptsGrpl %>% 
                                filter(IDpts == PointTree$IDpts[1])  %>%  
                                st_union() %>% 
                                st_cast("POINT") %>% 
                                st_as_sf() %>%
                                mutate(type = "Overlay") %>% 
                                mutate(ptsAcc = PointTree$ptsAcc[1]) %>% 
                                mutate(IDpts = PointTree$IDpts[1]) %>% 
                                mutate(origins = PointTree$origins[1])

              }else{

            Grpl2CblFlag <- TRUE

            ptsWIP$Logged<- FALSE

            for (j in 1:dim(ptsWIP)[1]) {
              ptsWIP$Logged[j] <- prod(as.integer(PointTree$origins[1][[1]] %in% ptsWIP$origins[[j]])+1) != 1
            }

            ptsWIP[ptsWIP$Logged == TRUE,"Logged"] <- NA
            ptsWIP <- na.exclude(ptsWIP)

            ptsWIP <- ptsWIP %>% 
              dplyr::select(-Logged) %>% 
              arrange(desc(n.overlaps))

            TmpPtsWIP <- ptsCbl %>% 
                                filter(IDpts == PointTree$IDpts[1])  %>%  
                                st_union() %>% 
                                st_cast("POINT") %>% 
                                st_as_sf() %>%
                                mutate(type = "Overlay") %>% 
                                mutate(ptsAcc = PointTree$ptsAcc[1]) %>% 
                                mutate(IDpts = PointTree$IDpts[1]) %>% 
                                mutate(origins = PointTree$origins[1])
              }

      }else{

        TmpPtsWIP <- ptsCbl %>% 
                                filter(IDpts == PointTree$IDpts[1])  %>%  
                                st_union() %>% 
                                st_cast("POINT") %>% 
                                st_as_sf() %>%
                                mutate(type = "Overlay") %>% 
                                mutate(ptsAcc = PointTree$ptsAcc[1]) %>% 
                                mutate(IDpts = PointTree$IDpts[1]) %>% 
                                mutate(origins = PointTree$origins[1])

      }

              TmpPtsWIP <- TmpPtsWIP  %>%  
                                filter(st_intersects(TmpPtsWIP,AccessMainTrails %>% st_union(),sparse = FALSE))

              TmpPtsWIP <- rbind(PointAcc %>%  
                                st_union() %>% 
                                st_cast("POINT")%>% 
                                st_as_sf() %>% 
                                mutate(type = "Access") %>% 
                                mutate(ptsAcc = NA ) %>% 
                                mutate(IDpts = NA) %>% 
                                mutate(origins = NA),TmpPtsWIP)

            CostDistEstWIP <-  AdjTopoLCP(costSurface = CondSurf,DEM = DEMmean , pts = TmpPtsWIP %>% as_Spatial(),
                              slopeRdCond = SlopeCond,paths = FALSE)[,1:dim(PointAcc)[1]]

            CostDistEstWIP <- CostDistEstWIP[(dim(PointAcc)[1]+1):dim(CostDistEstWIP)[1],]

          if (dim(PointAcc)[1] == 1) {
            PointTreeWIP <- TmpPtsWIP %>% 
              filter(type == "Overlay") %>% 
              mutate(EstCost = CostDistEstWIP) %>% 
              arrange(EstCost)
          }else{

            PointTreeWIP <- TmpPtsWIP %>% 
                filter(type == "Overlay") %>% 
                mutate(EstCost = apply(CostDistEstWIP,1, min)) %>% 
                arrange(EstCost)

          }

            TmpPtsWIP <- rbind(TmpPtsWIP %>% filter(type == "Access") %>% mutate(EstCost = NA),PointTreeWIP[1,])

            TmpPathWIP <- AdjTopoLCP(costSurface = CondSurf,DEM = DEMmean , pts = TmpPtsWIP %>% as_Spatial(),
                              slopeRdCond = SlopeCond,paths = FALSE)

            TmpPathWIPCost <- TmpPathWIP[1:dim(PointAcc)[1],dim(PointAcc)[1]+1]

            LCPathWIP <- max.col(t(-TmpPathWIPCost))

            TmpPtsWIP <- rbind(TmpPtsWIP[LCPathWIP,],PointTreeWIP[1,])

            TmpPathWIP <- AdjTopoLCP(costSurface = CondSurf,DEM = DEMmean , pts = TmpPtsWIP %>% as_Spatial(),
                              slopeRdCond = SlopeCond,paths = TRUE)

            CostRasterMean  <- rasterize(x = TmpPathWIP[[2]] ,
                                    y = CostRasterMean  , 
                                    field = 0.3,update =  TRUE)

            pathLines[[k]] <- TmpPathWIP[[2]]
            pathLines[[k]]@lines[[1]]@ID <- paste("Path",
                                                  LCPathWIP,
                                                  "-",
                                                  paste(as.character(unlist(PointTreeWIP$origins[1][[1]])), 
                                                        collapse='.'), 
                                                  sep = " ")

             k <- k +1

             if (Grpl2CblFlag) {

               ptsWIPCbl$Logged<- FALSE
              for (j in 1:dim(ptsWIPCbl)[1]) {
              ptsWIPCbl$Logged[j] <- prod(as.integer(PointTreeWIP$origins[1][[1]] %in% ptsWIPCbl$origins[[j]])+1) != 1
              }
                          ptsWIPCbl[ptsWIPCbl$Logged == TRUE,"Logged"] <- NA
            ptsWIPCbl <- na.exclude(ptsWIPCbl)
                        ptsWIPCbl <- ptsWIPCbl %>% 
              dplyr::select(-Logged) %>% 
              arrange(desc(n.overlaps))
             }else{

            ptsWIP$Logged<- FALSE

            for (j in 1:dim(ptsWIP)[1]) {
              ptsWIP$Logged[j] <- prod(as.integer(PointTreeWIP$origins[1][[1]] %in% ptsWIP$origins[[j]])+1) != 1
            }

                        ptsWIP[ptsWIP$Logged == TRUE,"Logged"] <- NA
            ptsWIP <- na.exclude(ptsWIP)

            ptsWIP <- ptsWIP %>% 
              dplyr::select(-Logged) %>% 
              arrange(desc(n.overlaps))

            if (winching == "2") {

              ptsWIPCbl$Logged<- FALSE
              for (j in 1:dim(ptsWIPCbl)[1]) {
              ptsWIPCbl$Logged[j] <- prod(as.integer(PointTreeWIP$origins[1][[1]] %in% ptsWIPCbl$origins[[j]])+1) != 1
              }
                          ptsWIPCbl[ptsWIPCbl$Logged == TRUE,"Logged"] <- NA
            ptsWIPCbl <- na.exclude(ptsWIPCbl)
                        ptsWIPCbl <- ptsWIPCbl %>% 
              dplyr::select(-Logged) %>% 
              arrange(desc(n.overlaps))
            }

            }

            if (dim(ptsWIP %>% filter(type == "Tree"))[1] == 0 & WinchingSwitch == FALSE) {
              winching <- "1"
              ptsWIP <- ptsWIPCbl
              WinchingSwitch <- TRUE
            }

            if (winching == "2") {
              RemainTree <- dim(ptsWIP %>% filter(type == "Tree"))[1] + dim(ptsWIPCbl %>% filter(type == "Tree"))[1]
            }else{
             RemainTree <- dim(ptsWIP %>% filter(type == "Tree"))[1] 
            }

    }

  }

```

```{r}
  paths <- do.call(rbind, pathLines)

    df <-  as_tibble(rasterToPoints(CostRasterMeanGrpl)) %>% mutate( Cost = layer) %>% dplyr::select(-layer)
  df2 <- as_tibble(rasterToPoints(DEM)) %>% mutate( z = Topo_P6_PARACOU) %>% dplyr::select(-Topo_P6_PARACOU)
  df3 <- as_tibble(rasterToPoints(raster(SlopeCond))) %>% mutate( AccSlp = layer) %>% dplyr::select(-layer)
  ggplot() +
    geom_raster(data = df, aes(x =x , y = y, fill = Cost)) +  geom_sf(data = pts) + scale_fill_gradientn(colours = rev(terrain.colors(225)))  + geom_sf( data = sf::st_as_sf(paths))    + geom_sf(data =  paths %>% st_as_sf()  %>% smoothr::smooth(method = "ksmooth",smoothness = 5) %>% st_buffer(dist = 4) %>% st_union(), color = 'red', alpha = 0.1 )  + theme_bw()
```