### 2021-08-25 R/Volume reviews functions.Rmd
loggingsummary and loggingsimulation functions

---
title: "Volume reviews functions"
author: "Vincyane Badouard"
date: "12/07/2021"
output: html_document
---

# timberharvestedvolume
INPUT : inventory + fuel
OUTPUT : TimberLoggedVolume

Si scenario != "RIL3fuelhollow"  ou (scenario == "manual"& fuel !="2"): 
TimberLoggedVolume =∑TreeHarvestableVolume  des arbres codés "Logged"= "1"

Si scenario = "RIL3fuelhollow" ou (scenario == "manual"& fuel =="2"): NoHollowTimberLoggedVolume = ∑TreeHarvestableVolume  des arbres codés "Logged"= "1"
TimberLoggedVolume = ∑NoHollowTimberLoggedVolume  et ⅔ du volume de ceux codés ProbedHollow = "1".

```{r}

LoggedTable <- inventory %>%
    filter(Logged == "1")

if (fuel !="2") {                    # no hollow trees exploitation

  TimberLoggedVolume <- sum(LoggedTable$TreeHarvestableVolume)
}

if (fuel =="2") {                   # with hollow trees exploitation

  HollowTable <- inventory %>%
    filter(ProbedHollow = "1")

  NoHollowTimberLoggedVolume <- sum(LoggedTable$TreeHarvestableVolume)
  TimberLoggedVolume <- sum(NoHollowTimberLoggedVolume +
                        1-(advancedloggingparameters$TreeHollowPartForFuel)*(HollowTable$TreeHarvestableVolume))
}

```

# exploitablefuelwoodvolume
*si (fuel =="1" |fuel =="2"))*

INPUT : inventory + TimberLoggedVolume/NoHollowTimberLoggedVolume
OUTPUT : FuelVolume

si scenario = "RIL3fuel" ou (scenario == "manual"& fuel =="1"):
FuelVolume = ∑TreeHarvestableVolume  de bille des arbres codés "maintrail", "2ndtrail", "treefall2nd" ou "landing" dans la colonne "DeathCause",  0,14*TimberLoggedVolume

si scenario = "RIL3fuelhollow" ou (scenario == "manual"& fuel =="2"):
FuelVolume = ∑TreeHarvestableVolume  de bille des arbres codés "maintrail", "2ndtrail", "treefall2nd" ou "landing" dans la colonne "DeathCause", ⅓ du volume de la bille des arbres codés "hollowfuel" dans la colonne "DeathCause",  0,14*NoHollowTimberLoggedVolume

```{r}

DamageTable <- inventory %>%
    filter(DeathCause = "maintrail" | DeathCause ="2ndtrail" | DeathCause = "treefall2nd" | DeathCause = "landing")

if (fuel =="0"){

DamageVolume <- sum(DamageTable$TreeHarvestableVolume) # Damage volume, non used for fuel in this scenario

}

if (fuel =="1"){

FuelVolume <- sum(DamageTable$TreeHarvestableVolume + advancedloggingparameters$Purge*TimberLoggedVolume) # Damage and purge (0.14)

}

if (fuel =="2"){

  HollowTable <- inventory %>%
    filter(DeathCause = "hollowfuel")

FuelVolume <- sum(DamageTable$TreeHarvestableVolume + 
                    advancedloggingparameters$TreeHollowPartForFuel *(HollowTable$TreeHarvestableVolume) + advancedloggingparameters$Purge*NoHollowTimberLoggedVolume)
# damage + 1/3 of hollow trees volume + purge (0.14) of the healthy trees

}
```