### 2021-12-13 R/AvoidAvRes_withcrowns.Rmd
Treefelling work

---
title: 'AvoidAvRes_withcrowns'
author: "Vincyane Badouard"
date: "08/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Maria)
library(sf)
library(dplyr)
library(nngeo)
library(ggplot2)
```

# Steps before the treefeling function
```{r}
MainTrail <- sf::st_linestring(matrix(c(286400, 582945,
                                        286400, 583250,
                                        286700, 583250,
                                        286700, 582945,
                                        286400, 582945) # the return
                                      ,ncol=2, byrow=TRUE))

pol1 <- list(matrix(c(286503, 583134,
                      286503, 583240,
                      286507, 583240,
                      286507, 583134,
                      286503, 583134) # the return
                    ,ncol=2, byrow=TRUE))
pol2 <- list(matrix(c(286650, 583134,
                      286650, 583240,
                      286654, 583240,
                      286654, 583134,
                      286650, 583134) # the return
                    ,ncol=2, byrow=TRUE))

PolList = list(pol1,pol2) #list of lists of numeric matrices
ScndTrail <- sf::st_multipolygon(PolList)

inventory <- addtreedim(inventorycheckformat(Paracou6_2016),
                        volumeparameters = ForestZoneVolumeParametersTable)

inventory <- suppressMessages(treeselection(inventory, objective = 20, scenario ="manual",
                                            fuel = "2", diversification = TRUE, specieslax = FALSE,
                                            objectivelax = TRUE, topography = DTMParacou, plotslope = PlotSlope,
                                            speciescriteria = SpeciesCriteria,
                                            advancedloggingparameters = loggingparameters())$inventory)

FutureReserveCrowns <- inventory %>% # create an object with future/reserve crowns only
  dplyr::filter(LoggingStatus == "future" | LoggingStatus == "reserve") %>%
  createcanopy() %>% # create all inventory crowns in the 'Crowns' column
  getgeometry(Crowns)

inventory <- inventory %>%
  dplyr::filter(Selected == "1") %>%
  dplyr::select(idTree,DBH,TrunkHeight,TreeHeight,CrownHeight,
                CrownDiameter,Selected, Xutm, Yutm)

dat <- inventory[2,] %>% # just 1 row (1 tree)
  # force the orientation success for the exemple
  tibble::add_column(TreeFellingOrientationSuccess = "1")
```

# Arguments values
```{r}
advancedloggingparameters = loggingparameters()
directionalfelling = "1"
fuel ="0"

```

# Preliminary steps
```{r}
# Function
Trail <- st_union(MainTrail, ScndTrail) # Our trail will be MainTrail or ScndTrail

# The crown
Crown <- dat %>%
  mutate(xCrown = Xutm,
         yCrown = Yutm + TrunkHeight + CrownHeight/2,
         exCrown = CrownDiameter/2,
         eyCrown = CrownHeight/2) %>%
  st_as_sf(coords = c("xCrown", "yCrown")) # ellipse centroid coordinates
Crown <- st_ellipse(Crown, Crown$exCrown, Crown$eyCrown) # create the ellipse

# The trunk
Trunk <- with(dat,
              st_polygon(list(matrix(c(Xutm-(DBH/100)/2, Yutm,
                                       Xutm-(DBH/100)/2, Yutm + TrunkHeight,
                                       Xutm+(DBH/100)/2, Yutm + TrunkHeight,
                                       Xutm+(DBH/100)/2, Yutm,
                                       Xutm-(DBH/100)/2, Yutm) # the return
                                     ,ncol=2, byrow=TRUE))))

# Find the point (TrailPt) on the Trail closest to the location of the tree (Foot)
Foot <- st_point(c(dat$Xutm,dat$Yutm)) # tree foot point

NearestPoints <- st_nearest_points(Foot, Trail) # from the Foot of the tree to the Trail (linestring)

NearestPoint <- st_cast(NearestPoints, "POINT") # to have start (Foot) and end (TrailPt) points
TrailPt <- NearestPoint[[2]] # the point (TrailPt) on the Trail closest to the location of the tree (Foot)

# Compute the angle between the tree default position and the shortest way from the foot to the trail
## Right-hand trail
if(TrailPt[1] >= Foot[1]){ # x trail > x foot

  theta <- as.numeric(matlib::angle(c(Foot[1] - Foot[1], dat$TreeHeight),
                                    c(TrailPt[1] - Foot[1], TrailPt[2] - Foot[2]), degree = TRUE))

  ## Left-hand trail
}else if(TrailPt[1] < Foot[1]){ # x trail < x foot

  theta <- 360 - numeric(matlib::angle(c(Foot[1] - Foot[1], dat$TreeHeight),
                                       c(TrailPt[1] - Foot[1], TrailPt[2] - Foot[2]), degree = TRUE))
}

## when the tree is on the trail (no angle)
if(is.na(theta)) theta <- 0

```

# Create future/reserve points
## Loop version
```{r}
# ## To direct to !avoid damage to future and reserve trees!, and foot directed to the trail
# # if (directionalfelling == "1"&& (fuel !="1" || fuel !="2")) {
# 
# # Fut/Res crowns/circumference as points
# FutResAllPoints <- list()
# for (FR in 1:nrow(FutureReserveCrowns)) {
#   ## if the fut/res tree is smaller or of the same height than the tree to be felled: we have to avoid the crown too
#   if(FutureReserveCrowns$TreeHeight[FR] <= dat$TreeHeight){
#     
#     FutResAllPoints[[FR]] <- st_cast(st_sfc(FutureReserveCrowns$Crowns[FR]), "POINT", group_or_split = TRUE) # crowns edge in points form
#     
#     ## if the fut/res tree is taller than the tree to be felled: we will only avoid it on its DBH
#   }else if(FutureReserveCrowns$TreeHeight[FR] > dat$TreeHeight){
#     
#     # PROBLEMES de list 
#     FutResAllPoints[[FR]] <- st_point(c(FutureReserveCrowns$Xutm[FR], FutureReserveCrowns$Yutm[FR])) %>%  # the fut/res tree foot point
#       st_buffer(dist = FutureReserveCrowns$DBH[FR]/2) %>%  # create a buffer of DBH diameter
#       st_cast("MULTIPOINT") %>%  # polygon to multipoint
#       st_sfc() %>% 
#       st_cast("POINT", group_or_split = TRUE)
#   }
# }
# # Get all these points back
# do.call(st_union, FutResAllPoints)

```

## On the data.frame version
```{r}
# if (directionalfelling != "0")

### Compute points
FutResCrownsPoints <- NULL # to start
FutResDBHPoints <- NULL # to start

# if the fut/res tree is smaller or of the same height than the tree to be felled: we have to avoid the crown too
if(any(FutureReserveCrowns$TreeHeight <= dat$TreeHeight)){

  FutResCrownsPoints <- FutureReserveCrowns %>% 
    filter(TreeHeight <= dat$TreeHeight) %>% 
    mutate(Crowns = st_cast(Crowns, "MULTIPOINT")) %>% # polygon to multipoint
    summarise(Points = st_combine(Crowns)) # several multipoints to 1
}

# if the fut/res tree is taller than the tree to be felled: we will only avoid it on its DBH
if(any(FutureReserveCrowns$TreeHeight > dat$TreeHeight)){

  FutResDBHPoints <- FutureReserveCrowns %>% 
    filter(TreeHeight > dat$TreeHeight) %>% 
    rowwise() %>% 
    mutate(Points = st_point(c(Xutm, Yutm)) %>% 
             st_buffer(dist = DBH/2) %>% 
             st_cast("MULTIPOINT") %>% 
             st_sfc()) %>% 
    ungroup() %>% 
    summarise(Points = st_combine(Points)) # several multipoints to 1
}

# Get all these points back
if(!is.null(FutResCrownsPoints) & !is.null(FutResDBHPoints)){ # # if smaller and taller trees
  FutResAllPoints <- bind_rows(FutResCrownsPoints, FutResDBHPoints) %>%
    summarise(Points = st_combine(Points)) # 2 multipoints to 1
}

if(!is.null(FutResCrownsPoints) & is.null(FutResDBHPoints)){ # if smaller trees but no tallers
  FutResAllPoints <- FutResCrownsPoints
}

if(!is.null(FutResDBHPoints) & is.null(FutResCrownsPoints)){ # if taller trees but no smallers
  FutResAllPoints <- FutResDBHPoints
}

# Multipoints to points
FutResAllPoints <- st_cast(st_sfc(FutResAllPoints$Points), "POINT", group_or_split = TRUE)

```
## Test time of each version
```{r}
# fun1 <- function(data, H = 30){
#   FutResAllPoints <- list()
#   for (FR in 1:nrow(data)) {
#     ## if the fut/res tree is smaller or of the same height than the tree to be felled: we have to avoid the crown too
#     if(data$TreeHeight[FR] <= H){
#       
#       FutResAllPoints[[FR]] <- st_cast(st_sfc(data$Crowns[FR]), "POINT", group_or_split = TRUE) # crowns edge in points form
#       
#       ## if the fut/res tree is taller than the tree to be felled: we will only avoid it on its DBH
#     }else if(data$TreeHeight[FR] > H){
#       
#       # PROBLEMES de list 
#       FutResAllPoints[[FR]] <- st_point(c(data$Xutm[FR], data$Yutm[FR])) %>%  # the fut/res tree foot point
#         st_buffer(dist = data$DBH[FR]/2) %>%  # create a buffer of DBH diameter
#         st_cast("MULTIPOINT") %>%  # polygon to multipoint
#         st_sfc() %>% 
#         st_cast("POINT", group_or_split = TRUE)
#     }
#   }
#   return(do.call(st_union, FutResAllPoints))
# }
# 
# fun2 <- function(data, H = 30){
#   
#   FutResCrownsPoints <- data %>% 
#     filter(TreeHeight <= H) %>% 
#     summarise(Points = st_combine(Crowns))
#   
#   FutResDBHPoints <- data %>% 
#     filter(TreeHeight > H) %>% 
#     rowwise() %>% 
#     mutate(Points = st_point(c(Xutm, Yutm)) %>% 
#              st_buffer(dist = DBH/2) %>% 
#              st_cast("MULTIPOINT") %>% # 
#              st_sfc()) %>% 
#     ungroup() %>% 
#     summarise(Points = st_combine(Points)) # several multipoints to 1
#   
#   FutResAllPoints <- bind_rows(FutResCrownsPoints, FutResDBHPoints) %>% 
#     summarise(Points = st_combine(Points))
#   
#   return(st_cast(st_sfc(FutResAllPoints$Points), "POINT", group_or_split = TRUE))
# }
# 
# microbenchmark::microbenchmark(fun1(FutureReserveCrowns), fun2(FutureReserveCrowns), times = 10)
```

# Distances and angles
```{r}

# Distances between each of these points and Foot
FutResDists <- rep(NA, length(FutResAllPoints)) # empty list for distances
RotationAngle <- rep(NA, length(FutResDists)) # empty list for rotation angles
OppAngToAvoid <- rep(NA, length(RotationAngle)) # empty list for angles to avoid
TrailAnglesToAvoid <- rep(NA, length(RotationAngle)) # empty list for angles to avoid
AnglesToAvoid <- rep(NA, length(RotationAngle)) # empty list for angles to avoid

for (FRP in 1:length(FutResAllPoints)) { # for each point
  FutResDists[[FRP]] <- st_distance(Foot, FutResAllPoints[[FRP]]) # compute the distance for each point

  # If distances < TreeHeight of the tree to be felled
  if(FutResDists[[FRP]] < dat$TreeHeight){
    # sum(FutResDists < dat$TreeHeight) # check av/res trees close to the tree to be felled

    # Compute the angles to the trail of this points
    ## 1st: the rotation angle (between the default position of the tree to cut down, and the position to avoid)
    ### Right-hand fut/res point
    if(FutResAllPoints[[FRP]][1] >= Foot[1]){ # x FRP > x foot

      RotationAngle[[FRP]] <- as.numeric(matlib::angle(
        c(Foot[1] - Foot[1], dat$TreeHeight), # default position
        c(FutResAllPoints[[FRP]][1] - Foot[1], FutResAllPoints[[FRP]][2] - Foot[2]), # position to avoid
        degree = TRUE))

      ### Left-hand fut/res point
    }else if(FutResAllPoints[[FRP]][1] < Foot[1]){ # x FRP < x foot

      RotationAngle[[FRP]] <- 360 - as.numeric(matlib::angle(
        c(Foot[1] - Foot[1], dat$TreeHeight), # default position
        c(FutResAllPoints[[FRP]][1] - Foot[1], FutResAllPoints[[FRP]][2] - Foot[2]), # position to avoid
        degree = TRUE))
      }

    # sum(!is.na(RotationAngle)) # check for how many av/res trees the angle has been computed

    ## Then the angle to the trail: the angle to avoid
    OppAngToAvoid[[FRP]] <- RotationAngle[[FRP]] - 180 - theta
    TrailAnglesToAvoid[[FRP]] <- 180-(90 + OppAngToAvoid[[FRP]])

    # Record their angles if there are in the fall zone ([1; 179°])
    if(TrailAnglesToAvoid[[FRP]] > 1 & TrailAnglesToAvoid[[FRP]] < 179){
      # any(!is.na(TrailAnglesToAvoid) < 179)

      AnglesToAvoid[[FRP]] <- round(TrailAnglesToAvoid[[FRP]], digits = 0)
      # Plus1 <- AnglesToAvoid + 1
      # Moins1 <- AnglesToAvoid - 1
      # Plus2 <- AnglesToAvoid + 2
      # Moins2 <- AnglesToAvoid - 2 # to add a 2° buffer
      # AnglesToAvoid <- c(AnglesToAvoid, Plus1, Moins1) #, Plus2, Moins2
      # sum(!is.na(Plus1)) to check

    }else{AnglesToAvoid <- NULL} # No angles to avoid

    # sum(!is.na(AnglesToAvoid)) # check for how many av/res trees the angle has been computed
  }else{AnglesToAvoid <- NULL} # No angles to avoid
}

```
# Tree felling
```{r}
# Drop the tree between [1; 179°] avoiding the angles of the futures/reserves
if(dat$TreeFellingOrientationSuccess == "1"){

  TreefallOrientation <- c(1:179) # the a priori fall zone (angles to the trail)
  TreefallOrientation <- TreefallOrientation[!TreefallOrientation %in% unique(AnglesToAvoid)] # Remove angles (the fut/res trees) to avoid

  TreefallOrientation <- as.numeric(sample(TreefallOrientation, size = 1)) # only those angles that avoid the fut/res trees

  OppAng <- 180-(90 + TreefallOrientation) # the angle between the closest position to the trail (90°) and the desired position (desired angle to the trail)

  ## Right-hand trail
  if(TrailPt[1] >= Foot[1]){ # x trail > x foot

    Aangle <- round(as.numeric(180 + OppAng + theta), digits = 0) # Foot oriented

    ## Left-hand trail
  }else if(TrailPt[1] < Foot[1]){ # x trail < x foot

    Aangle <- round(as.numeric(360 - OppAng + theta), digits = 0) # Foot oriented
  }

  # Calculate the crown and trunk position
  FallenTree <- st_difference(st_union( # A configuration (90° to the trail (shortest))
    rotatepolygon(Trunk, angle = Aangle, fixed = Foot), # turned trunk
    rotatepolygon(Crown, angle = Aangle, fixed = Foot) # turned crown
  ))

}else{ # else random felling
  RandomAngle <- as.numeric(sample(c(0:359.9), size = 1))

  FallenTree <- st_difference(st_union(
    rotatepolygon(Trunk, angle = RandomAngle, fixed = Foot), # turned trunk
    rotatepolygon(Crown, angle = RandomAngle, fixed = Foot) # turned crown
  ))
}

```

```{r}
ggplot() +
  geom_sf(data = Foot) +
  geom_sf(data = Trail) +
  geom_sf(data = NearestPoints) +
  geom_sf(data = TrailPt) +
  geom_sf(data = FallenTree) +
  geom_sf(data = FutureReserveCrowns) +
  geom_sf(data = FutResCrownsPoints, col = "green") + # smaller trees
  geom_sf(data = FutResDBHPoints, col = "pink")

# geom_sf(data = FutResAllPoints, col = "green")
```