### 2021-12-13 R/AvoidAvRes_withoutcrowns.Rmd

---
title: "AvoidAvRes_withoutcrowns"
author: "Vincyane Badouard"
date: "06/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Maria)
library(sf)
library(dplyr)
library(nngeo)
library(ggplot2)
```

```{r}
MainTrail <- sf::st_linestring(matrix(c(286400, 582945,
                                        286400, 583250,
                                        286700, 583250,
                                        286700, 582945,
                                        286400, 582945) # the return
                                      ,ncol=2, byrow=TRUE))

pol1 <- list(matrix(c(286503, 583134,
                      286503, 583240,
                      286507, 583240,
                      286507, 583134,
                      286503, 583134) # the return
                    ,ncol=2, byrow=TRUE))
pol2 <- list(matrix(c(286650, 583134,
                      286650, 583240,
                      286654, 583240,
                      286654, 583134,
                      286650, 583134) # the return
                    ,ncol=2, byrow=TRUE))

PolList = list(pol1,pol2) #list of lists of numeric matrices
ScndTrail <- sf::st_multipolygon(PolList)

inventory <- addtreedim(inventorycheckformat(Paracou6_2016),
                        volumeparameters = ForestZoneVolumeParametersTable)

inventory <- suppressMessages(treeselection(inventory, objective = 20, scenario ="manual",
                                            fuel = "2", diversification = TRUE, specieslax = FALSE,
                                            objectivelax = TRUE, topography = DTMParacou, plotslope = PlotSlope,
                                            speciescriteria = SpeciesCriteria,
                                            advancedloggingparameters = loggingparameters())$inventory)

FutureReserveCrowns <- inventory %>% # create an object with future/reserve crowns only
  dplyr::filter(LoggingStatus == "future" | LoggingStatus == "reserve") %>%
  createcanopy() %>% # create all inventory crowns in the 'Crowns' column
  getgeometry(Crowns)

inventory <- inventory %>%
  dplyr::filter(Selected == "1") %>%
  dplyr::select(idTree,DBH,TrunkHeight,TreeHeight,CrownHeight,
                CrownDiameter,Selected, Xutm, Yutm)

dat <- inventory[1,] %>% # just 1 row (1 tree)
  # force the orientation success for the exemple
  tibble::add_column(TreeFellingOrientationSuccess = "1")
```

```{r}
FutureReserveCrowns$Crowns <- st_cast(FutureReserveCrowns$Crowns, "MULTIPOINT")

ggplot() +
   geom_sf(data = getgeometry(FutureReserveCrowns, Crowns), 
           col = "forestgreen")
plot(FutResCrownsPoints)
```

# Arguments values
```{r}
advancedloggingparameters = loggingparameters()
directionalfelling = "1"
fuel ="0"

```

Preliminary steps
```{r}
# Function
Trail <- st_union(MainTrail, ScndTrail) # Our trail will be MainTrail or ScndTrail

# The crown
Crown <- dat %>%
  mutate(xCrown = Xutm,
         yCrown = Yutm + TrunkHeight + CrownHeight/2,
         exCrown = CrownDiameter/2,
         eyCrown = CrownHeight/2) %>%
  st_as_sf(coords = c("xCrown", "yCrown")) # ellipse centroid coordinates
Crown <- st_ellipse(Crown, Crown$exCrown, Crown$eyCrown) # create the ellipse

# The trunk
Trunk <- with(dat,
              st_polygon(list(matrix(c(Xutm-(DBH/100)/2, Yutm,
                                       Xutm-(DBH/100)/2, Yutm + TrunkHeight,
                                       Xutm+(DBH/100)/2, Yutm + TrunkHeight,
                                       Xutm+(DBH/100)/2, Yutm,
                                       Xutm-(DBH/100)/2, Yutm) # the return
                                     ,ncol=2, byrow=TRUE))))

# Find the point (TrailPt) on the Trail closest to the location of the tree (Foot)
Foot <- st_point(c(dat$Xutm,dat$Yutm)) # tree foot point

NearestPoints <- st_nearest_points(Foot, Trail) # from the Foot of the tree to the Trail (linestring)

NearestPoint <- st_cast(NearestPoints, "POINT") # to have start (Foot) and end (TrailPt) points
TrailPt <- NearestPoint[[2]] # the point (TrailPt) on the Trail closest to the location of the tree (Foot)

# Compute the angle between the tree default position and the shortest way from the foot to the trail
theta <- as.numeric(matlib::angle(c(Foot[1] - Foot[1], dat$TreeHeight),
                                  c(TrailPt[1] - Foot[1], TrailPt[2] - Foot[2]), degree = TRUE))

## when the tree is on the trail (no angle)
if(is.na(theta)) theta <- 0

```

New work
```{r}
## To direct to !avoid damage to future and reserve trees!, and foot directed to the trail
# if (directionalfelling == "1"&& (fuel !="1" || fuel !="2")) {

# Find the future/reserve on the fall zone
FutRescoord <-  FutureReserveCrowns %>%
  filter(LoggingStatus == "future"| LoggingStatus == "reserve") %>%
  select(Xutm, Yutm) 
st_geometry(FutRescoord) <- NULL

FutResFoots <- st_multipoint(x = as.matrix(FutRescoord)) # all the future/reserve trees foot
FutResFoots <- st_cast(st_sfc(FutResFoots), "POINT", group_or_split = TRUE) # multiple points

## distance to Foot < TreeHeight
FutResDists <- rep(NA, length(FutResFoots)) # empty list for distances
RotationAngle <- rep(NA, length(FutResDists)) # empty list for rotation angles
OppAngToAvoid <- rep(NA, length(RotationAngle)) # empty list for angles to avoid
TrailAnglesToAvoid <- rep(NA, length(RotationAngle)) # empty list for angles to avoid
AnglesToAvoid <- rep(NA, length(RotationAngle)) # empty list for angles to avoid

for (FRF in 1:length(FutResFoots)) { # for each point
  FutResDists[[FRF]] <- st_distance(Foot, FutResFoots[[FRF]]) # compute the distance for each point

  ## angle to the trail
  if(FutResDists[[FRF]] < dat$TreeHeight){
    # sum(FutResDists < dat$TreeHeight) # check av/res trees close to the tree to be felled

    ### 1st: the rotation angle (between the default position of the tree to cut down, and the position to avoid)
    RotationAngle[[FRF]] <- as.numeric(matlib::angle(
      c(Foot[1] - Foot[1], dat$TreeHeight), # default position
      c(FutResFoots[[FRF]][1] - Foot[1], FutResFoots[[FRF]][2] - Foot[2]), # position to avoid
      degree = TRUE))
    # sum(!is.na(RotationAngle)) # check for how many av/res trees the angle has been computed

    ### Then the angle to the trail: the angle to avoid
    OppAngToAvoid[[FRF]] <- RotationAngle[[FRF]] - 180 - theta
    TrailAnglesToAvoid[[FRF]] <- 180-(90 + OppAngToAvoid[[FRF]])

    ## record their angles if there are in the fall zone ([0.1; 179.9°])
    if(TrailAnglesToAvoid[[FRF]] %in% c(0.1:179.9)){

      AnglesToAvoid[[FRF]] <- TrailAnglesToAvoid[[FRF]]
    }else{AnglesToAvoid <- NULL} # No angles to avoid

  }} # temporaires

```

```{r}
# Drop the tree between [0.1; 179.9°] avoiding the angles of the futures/reserves
if(dat$TreeFellingOrientationSuccess == "1"){

  TreefallOrientation <- c(1:179) # the a priori fall zone (angles to the trail)
  TreefallOrientation <- TreefallOrientation[!TreefallOrientation %in% AnglesToAvoid] # Remove angles (the fut/res trees) to avoid

  TreefallOrientation <- as.numeric(sample(TreefallOrientation, size = 1)) # only those angles that avoid the fut/res trees

  OppAng <- 180-(90 + TreefallOrientation) # the angle between the closest position to the trail (90°) and the desired position (desired angle to the trail)

  ## Right-hand trail
  if(TrailPt[1] >= Foot[1]){ # x trail > x foot

    Aangle <- round(as.numeric(180 + OppAng + theta), digits = 0) # Foot oriented

    ## Left-hand trail
  }else if(TrailPt[1] < Foot[1]){ # x trail < x foot

    Aangle <- round(as.numeric(360 - OppAng + theta), digits = 0) # Foot oriented
  }

  # Calculate the crown and trunk position
  FallenTree <- st_difference(st_union( # A configuration (90° to the trail (shortest))
    rotatepolygon(Trunk, angle = Aangle, fixed = Foot), # turned trunk
    rotatepolygon(Crown, angle = Aangle, fixed = Foot) # turned crown
  ))

}else{ # else random felling
  RandomAngle <- as.numeric(sample(c(0:359.9), size = 1))

  FallenTree <- st_difference(st_union(
    rotatepolygon(Trunk, angle = RandomAngle, fixed = Foot), # turned trunk
    rotatepolygon(Crown, angle = RandomAngle, fixed = Foot) # turned crown
  ))
}
```