### 2021-12-16 R/AvoidFutRessimpletest.Rmd
Treefelling end!!

---
title: "AvoidFutRessimpletest"
author: "Vincyane Badouard"
date: "10/12/2021"
output: html_document
---
```{r}
rm(list = ls()) ; invisible(gc())
library(Maria)
library(sf)
library(dplyr)
library(nngeo)
library(ggplot2)
set.seed(420)
```

# Trails

```{r}
MainTrail <- sf::st_linestring(matrix(c(286400, 582945,
                                        286400, 583250,
                                        286700, 583250,
                                        286700, 582945,
                                        286400, 582945) # the return
                                      ,ncol=2, byrow=TRUE))
pol1 <- list(matrix(c(286503, 583134,
                      286503, 583240,
                      286507, 583240,
                      286507, 583134,
                      286503, 583134) # the return
                    ,ncol=2, byrow=TRUE))
pol2 <- list(matrix(c(286650, 583134,
                      286650, 583240,
                      286654, 583240,
                      286654, 583134,
                      286650, 583134) # the return
                    ,ncol=2, byrow=TRUE))

PolList = list(pol1,pol2) #list of lists of numeric matrices
ScndTrail <- sf::st_multipolygon(PolList)
Trail <- st_union(MainTrail, ScndTrail) # Our trail will be MainTrail or ScndTrail
rm(MainTrail, ScndTrail, pol1, pol2, PolList)
```

```{r}
ggplot() +
  geom_sf(data = Trail)
```

# Individuals

```{r}
inventory <- addtreedim(inventorycheckformat(Paracou6_2016),
                        volumeparameters = ForestZoneVolumeParametersTable)

inventory <- suppressMessages(treeselection(inventory, objective = 20, scenario ="manual",
                                            fuel = "2", diversification = TRUE, specieslax = FALSE,
                                            objectivelax = TRUE, topography = DTMParacou, 
                                            plotslope = PlotSlope,
                                            speciescriteria = SpeciesCriteria,
                                            advancedloggingparameters = loggingparameters())$inventory)
dat <- inventory %>%
  dplyr::filter(Selected == "1") %>%
  dplyr::select(idTree,DBH,TrunkHeight,TreeHeight,CrownHeight,
                CrownDiameter,Selected, Xutm, Yutm) %>% 
  sample_n(1) %>% 
  tibble::add_column(TreeFellingOrientationSuccess = "1")

dat$Xutm <- 286546.1
dat$Yutm <- 583029.1
Foot <- st_point(c(dat$Xutm,dat$Yutm))
```

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot)
```

# Future Reserve

```{r}
FutureReserveCrowns <- inventory %>% # create an object with future/reserve crowns only
  dplyr::filter(LoggingStatus == "future" | LoggingStatus == "reserve") %>%
  createcanopy() %>% # create all inventory crowns in the 'Crowns' column
  getgeometry(Crowns)

FutureReserveCrowns <- FutureReserveCrowns[1:2,]

# at the same y than the tree to be felled
# 1 on the right side
FutureReserveCrowns[1,]$Xutm <- dat$Xutm + 30
FutureReserveCrowns[1,]$Yutm <- dat$Yutm + 10

# 1 on the left side
FutureReserveCrowns[2,]$Xutm <- dat$Xutm - 30 
FutureReserveCrowns[2,]$Yutm <- dat$Yutm + 10

# # Re-compute the crown
st_geometry(FutureReserveCrowns) <- NULL

FutureReserveCrowns <- FutureReserveCrowns %>%
  createcanopy() %>% # create all inventory crowns in the 'Crowns' column
  getgeometry(Crowns)

rm(inventory)
```

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot) +
  geom_sf(data = FutureReserveCrowns$Crowns, col = "green")
```

# Individual geometry

```{r}
# The crown
Crown <- dat %>%
  mutate(xCrown = Xutm,
         yCrown = Yutm + TrunkHeight + CrownHeight/2,
         exCrown = CrownDiameter/2,
         eyCrown = CrownHeight/2) %>%
  st_as_sf(coords = c("xCrown", "yCrown")) # ellipse centroid coordinates
Crown <- st_ellipse(Crown, Crown$exCrown, Crown$eyCrown) # create the ellipse

# The trunk
Trunk <- with(dat,
              st_polygon(list(matrix(c(Xutm-(DBH/100)/2, Yutm,
                                       Xutm-(DBH/100)/2, Yutm + TrunkHeight,
                                       Xutm+(DBH/100)/2, Yutm + TrunkHeight,
                                       Xutm+(DBH/100)/2, Yutm,
                                       Xutm-(DBH/100)/2, Yutm) # the return
                                     ,ncol=2, byrow=TRUE))))

# Find the point (TrailPt) on the Trail closest to the location of the tree (Foot)
# tree foot point

NearestPoints <- st_nearest_points(Foot, Trail) # from the Foot of the tree to the Trail (linestring)

NearestPoint <- st_cast(NearestPoints, "POINT") # to have start (Foot) and end (TrailPt) points
TrailPt <- NearestPoint[[2]] # the point (TrailPt) on the Trail closest to the location of the tree (Foot)
rm(NearestPoint)

# Compute the angle between the tree default position and the shortest way from the foot to the trail
## Right-hand trail
if(TrailPt[1] >= Foot[1]){ # x trail > x foot

  theta <- as.numeric(matlib::angle(c(Foot[1] - Foot[1], dat$TreeHeight),
                                    c(TrailPt[1] - Foot[1], TrailPt[2] - Foot[2]), degree = TRUE))

  ## Left-hand trail
}else if(TrailPt[1] < Foot[1]){ # x trail < x foot

  theta <- 180 - numeric(matlib::angle(c(Foot[1] - Foot[1], dat$TreeHeight),
                                       c(TrailPt[1] - Foot[1], TrailPt[2] - Foot[2]), degree = TRUE))
}

## when the tree is on the trail (no angle)
if(is.na(theta)) theta <- 0L
theta <- theta
```

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot) +
  geom_sf(data = FutureReserveCrowns$Crowns, col = "green") +
  geom_sf(data = NearestPoints) +
  geom_sf(data = TrailPt) +
  geom_sf(data = Crown) +
  geom_sf(data = Trunk)
```

# Future Reserve Crowns to Points

```{r}
# if (directionalfelling != "0")

### Compute points
FutResCrownsPoints <- NULL # to start
FutResDBHPoints <- NULL # to start

# if the fut/res tree is smaller or of the same height than the tree to be felled: we have to avoid the crown too
if(any(FutureReserveCrowns$TreeHeight <= dat$TreeHeight)){

  FutResCrownsPoints <- FutureReserveCrowns %>% 
    filter(TreeHeight <= dat$TreeHeight) %>% 
    mutate(Crowns = st_cast(Crowns, "MULTIPOINT")) %>% # polygon to multipoint
    summarise(Points = st_combine(Crowns)) # several multipoints to 1
}

# if the fut/res tree is taller than the tree to be felled: we will only avoid it on its DBH
if(any(FutureReserveCrowns$TreeHeight > dat$TreeHeight)){

  FutResDBHPoints <- FutureReserveCrowns %>% 
    filter(TreeHeight > dat$TreeHeight) %>% 
    rowwise() %>% 
    mutate(Points = st_point(c(Xutm, Yutm)) %>% 
             st_buffer(dist = DBH/2) %>% 
             st_cast("MULTIPOINT") %>% 
             st_sfc()) %>% 
    ungroup() %>% 
    summarise(Points = st_combine(Points)) # several multipoints to 1
}

# Get all these points back
if(!is.null(FutResCrownsPoints) & !is.null(FutResDBHPoints)){ # # if smaller and taller trees
  FutResAllPoints <- bind_rows(FutResCrownsPoints, FutResDBHPoints) %>%
    summarise(Points = st_combine(Points)) # 2 multipoints to 1
}

if(!is.null(FutResCrownsPoints) & is.null(FutResDBHPoints)){ # if smaller trees but no tallers
  FutResAllPoints <- FutResCrownsPoints
}

if(!is.null(FutResDBHPoints) & is.null(FutResCrownsPoints)){ # if taller trees but no smallers
  FutResAllPoints <- FutResDBHPoints
}

# Multipoints to points
FutResAllPoints <- st_cast(st_sfc(FutResAllPoints$Points), "POINT", group_or_split = TRUE)

rm(FutResCrownsPoints, FutResDBHPoints)
```

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot) +
  geom_sf(data = FutureReserveCrowns$Crowns, col = "green") +
  geom_sf(data = NearestPoints) +
  geom_sf(data = TrailPt) +
  geom_sf(data = Crown) +
  geom_sf(data = Trunk) +
  geom_sf(data = FutResAllPoints, col = "green")
```

# Individual Foot Buffer

```{r}
FootBuffer <- dat %>% 
  mutate(Points = st_point(c(Xutm, Yutm)) %>% 
           st_buffer(dist = CrownDiameter/2) %>% 
           st_cast("MULTIPOINT") %>% 
           st_sfc())
# Multipoints to points
FootBuffer <- st_cast(st_sfc(FootBuffer$Points), "POINT", group_or_split = TRUE)
```

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot) +
  geom_sf(data = FutureReserveCrowns$Crowns, col = "green") +
  geom_sf(data = NearestPoints) +
  geom_sf(data = TrailPt) +
  geom_sf(data = Crown) +
  geom_sf(data = Trunk) +
  geom_sf(data = FutResAllPoints, col = "green") +
  geom_sf(data = FootBuffer)
```

# Distances and angles
```{r}
# Only Foots 
# FutResAllPoints <- FutureReserveCrowns %>% 
#    select(Xutm, Yutm)
# 
# FutResFoots <- st_multipoint(x = as.matrix(FutResAllPoints)) # all the future/reserve trees foot
# FutResAllPoints <- st_cast(st_sfc(FutResFoots), "POINT", group_or_split = TRUE) # multiple points

# Distances between each of these points and Foot
RotationAngle <- c() # empty vector for rotation angles
for (FRP in 1:length(FutResAllPoints)) { # for each point
  FutResDist <- st_distance(Foot, FutResAllPoints[[FRP]]) # compute the distance for each point

  # If distances < TreeHeight of the tree to be felled
  if(FutResDist < dat$TreeHeight){
    # sum(FutResDists < dat$TreeHeight) # check av/res trees close to the tree to be felled

    # Compute the angles to the trail of this points
    ## 1st: the rotation angle (between the default position of the tree to cut down, and the position to avoid)
    for (P in 1:length(FootBuffer)) { # for each point of the Foot buffer

      ### Right-hand fut/res point
      if(FutResAllPoints[[FRP]][1] >= FootBuffer[[P]][1]){ # x FRP > x foot

        RotationAngle <- c(RotationAngle, as.numeric(matlib::angle(
          c(FootBuffer[[P]][1] - FootBuffer[[P]][1], dat$TreeHeight), # default position
          c(FutResAllPoints[[FRP]][1] - FootBuffer[[P]][1], FutResAllPoints[[FRP]][2] - FootBuffer[[P]][2]), # position to avoid
          degree = TRUE)- 180))

        ### Left-hand fut/res point
      }else if(FutResAllPoints[[FRP]][1] < FootBuffer[[P]][1]){ # x FRP < x foot

        RotationAngle <- c(RotationAngle, as.numeric(matlib::angle(
          c(FootBuffer[[P]][1] - FootBuffer[[P]][1], dat$TreeHeight), # default position
          c(FutResAllPoints[[FRP]][1] - FootBuffer[[P]][1], FutResAllPoints[[FRP]][2] - 
              FootBuffer[[P]][2]), # position to avoid
          degree = TRUE)))
      }
    }
  }else{AnglesToAvoid <- NULL}
}
rm(FutResDist)
# sum(!is.na(RotationAngle)) # check for how many av/res trees the angle has been computed

## Then the angle to the trail: the angle to avoid
OppAngToAvoid <- RotationAngle - theta

TrailAnglesToAvoid <- 180-(90 + OppAngToAvoid)
TrailAnglesToAvoid <- round(TrailAnglesToAvoid, digits = 0)
TrailAnglesToAvoid[TrailAnglesToAvoid >= 360] <- TrailAnglesToAvoid[TrailAnglesToAvoid >= 360] - 360

AnglesToAvoid <- TrailAnglesToAvoid[
 TrailAnglesToAvoid >= 1 & TrailAnglesToAvoid <= 179
]
```

# One tree felling

```{r}
# Drop the tree between [1; 179°] avoiding the angles of the futures/reserves
if(dat$TreeFellingOrientationSuccess == "1"){

  TreefallOrientation <- c(1:179) # the a priori fall zone (angles to the trail)
  TreefallOrientation <- TreefallOrientation[!TreefallOrientation %in% unique(AnglesToAvoid)] # Remove angles (the fut/res trees) to avoid

  TreefallOrientation <- as.numeric(sample(TreefallOrientation, size = 1)) # only those angles that avoid the fut/res trees

  OppAng <- 180-(90 + TreefallOrientation) # the angle between the closest position to the trail (90°) and the desired position (desired angle to the trail)

  ## Right-hand trail
  if(TrailPt[1] >= Foot[1]){ # x trail > x foot

    Aangle <- round(as.numeric(180 + OppAng + theta), digits = 0) # Foot oriented

    ## Left-hand trail
  }else if(TrailPt[1] < Foot[1]){ # x trail < x foot

    Aangle <- round(as.numeric(360 - OppAng + theta), digits = 0) # Foot oriented
  }

  # Calculate the crown and trunk position
  FallenTree <- st_difference(st_union( # A configuration (90° to the trail (shortest))
    rotatepolygon(Trunk, angle = Aangle, fixed = Foot), # turned trunk
    rotatepolygon(Crown, angle = Aangle, fixed = Foot) # turned crown
  ))

}else{ # else random felling
  RandomAngle <- as.numeric(sample(c(0:359), size = 1))

  FallenTree <- st_difference(st_union(
    rotatepolygon(Trunk, angle = RandomAngle, fixed = Foot), # turned trunk
    rotatepolygon(Crown, angle = RandomAngle, fixed = Foot) # turned crown
  ))
}
```

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot) +
  geom_sf(data = FutureReserveCrowns$Crowns, col = "green") +
  geom_sf(data = NearestPoints) +
  geom_sf(data = TrailPt) +
  geom_sf(data = Crown) +
  geom_sf(data = Trunk) +
  geom_sf(data = FutResAllPoints, col = "green") +
  geom_sf(data = FootBuffer) +
  geom_sf(data = FallenTree, col = "red")
```

# All tree felling

```{r}
TreefallOrientation <- c(1:179) # the a priori fall zone (angles to the trail)
TreefallOrientation <- TreefallOrientation[!TreefallOrientation %in% unique(AnglesToAvoid)] # Remove angles (the fut/res trees) to avoid

felling <- function(theta){

  OppAng <- 180-(90 + as.numeric(theta)) # the angle between the closest position to the trail (90°) and the desired position (desired angle to the trail)

  ## Right-hand trail
  if(TrailPt[1] >= Foot[1]){ # x trail > x foot

    Aangle <- round(as.numeric(180 + OppAng + theta), digits = 0) # Foot oriented

    ## Left-hand trail
  }else if(TrailPt[1] < Foot[1]){ # x trail < x foot

    Aangle <- round(as.numeric(360 - OppAng + theta), digits = 0) # Foot oriented
  }

  # Calculate the crown and trunk position
  FallenTree <- st_difference(st_union( # A configuration (90° to the trail (shortest))
    rotatepolygon(Trunk, angle = Aangle, fixed = Foot), # turned trunk
    rotatepolygon(Crown, angle = Aangle, fixed = Foot) # turned crown
  ))
  return(FallenTree)
}

FallenTrees <- lapply(as.list(TreefallOrientation), felling)
FallenTrees <- lapply(FallenTrees, function(x) data.frame(Tree = x)) %>% 
  bind_rows()
```

```{r}
ggplot() +
  geom_sf(data = Trail) +
  geom_sf(data = Foot) +
  geom_sf(data = FutureReserveCrowns$Crowns, col = "green") +
  geom_sf(data = NearestPoints) +
  geom_sf(data = TrailPt) +
  geom_sf(data = Crown) +
  geom_sf(data = Trunk) +
  geom_sf(data = FutResAllPoints, col = "green") +
  geom_sf(data = FootBuffer) +
  geom_sf(data = FallenTrees$geometry, col = "red")
```