### 2021-12-29 R/Remove trees out of the plot.Rmd

---
title: "Remove trees out of the plot"
author: "Vincyane Badouard"
date: "28/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(Maria)
library(ggplot2)
library(sf)
library(dplyr)
library(tibble)
```

# Necessary data
```{r}
data(Paracou6_2016)
data(PlotMask) # SpatialPolygonsDataFrame (sp)

PlotMask <- st_as_sf(PlotMask) # sp to sf (sfc_POLYGON)

ggplot() +
  geom_sf(data = PlotMask, col = "red")
# PlotMask is only the boundaries of the plot
```

# Remove trees out of the plot

```{r}
inventory <- st_as_sf(Paracou6_2016, coords = c("Xutm", "Yutm")) %>% 
  st_set_crs(st_crs(PlotMask)) # set crs

TreesIn <- suppressWarnings(st_intersection(inventory, PlotMask)) %>% 
  add_column(TreesIn = "1") %>%
  select(idTree, TreesIn)

st_geometry(TreesIn) <- NULL # sf to data.frame
TreesIn <- unique(TreesIn)

inventory <- inventory %>%
  left_join(TreesIn, by = "idTree") %>% 
   mutate(TreesIn = ifelse(is.na(TreesIn),
                               "0", TreesIn))
  # select(-TreesIn)

class(inventory)
```

# Plot in & out trees
```{r}
In <- sf::st_as_sf(
  filter(inventory, TreesIn == "1"), coords = c("Xutm", "Yutm"))

Out <- sf::st_as_sf(
  filter(inventory, TreesIn == "0"), coords = c("Xutm", "Yutm"))

ggplot() +
  geom_sf(data = In, col = "green") +
  geom_sf(data = Out, col = "red")
```