### 2021-12-30 R/Iterations summary.Rmd

---
title: "Iterations summary"
author: "Vincyane Badouard"
date: "29/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# save(Rslt, file = "../data/Rslt.Rdata")
load("../data/Rslt.Rdata")
```

Les stats qu'on veut faire : mean, sd, quantile

Sur les variables : c(HarvestableArea, VO, HVinit, TimberLoggedVolume, NoHollowTimberLoggedVolume, FuelVolume,
DamageVolume, LostBiomass)

# With 1 var
```{r}
HarvestableArea_mean <- mean(c(Rslt[[1]]$HarvestableArea, Rslt[[2]]$HarvestableArea))
HarvestableArea_sd <- sd(c(Rslt[[1]]$HarvestableArea, Rslt[[2]]$HarvestableArea))
HarvestableArea_quantile <- quantile(c(Rslt[[1]]$HarvestableArea, Rslt[[2]]$HarvestableArea))

```

# pour toutes les vars pour n iter
```{r}
var <- c("HarvestableArea", "VO", "HVinit", "TimberLoggedVolume", "NoHollowTimberLoggedVolume", "FuelVolume",
         "DamageVolume", "LostBiomass")

quantile(c(as.numeric(Rslt[[1]]["LostBiomass"]) : as.numeric(Rslt[[2]]["LostBiomass"])), probs = c(0.25, 0.75))

statsvars <- function(v, iter){

  itervalues <- c(as.numeric(Rslt[[1]][v]) : as.numeric(Rslt[[length(Rslt)]][v]))

  Mean <- round(mean(itervalues), digits = 1)
  Sd <- round(sd(itervalues), digits = 1)
  Q1 <- round(quantile(itervalues, probs = 0.25, digits = 1))
  Median <- round(median(itervalues), digits = 1)
  Q3 <- round(quantile(itervalues, probs = 0.75, digits = 1))
  Min <- round(min(itervalues), digits = 1)
  Max <- round(max(itervalues), digits = 1)

  output <- list(Mean = Mean, Sd = Sd, Min = Min, Q1 = Q1, Median = Median, Q3 = Q3, Max = Max)

  return(output)
}

RsltStats <- lapply(var, statsvars, 2)

unlist(RsltStats[8]) # numeric named vector

```