### 2023-08-30 docs/Article example.Rmd
try to fix 2nd trails and harvestableareadef issues

---
title: "Article example"
author: "Vincyane Badouard"
date: "`r Sys.Date()`"
output: html_document
---

# Data and libraries
```{r}
data(Paracou6_2016) # inventory
data(PlotMask) # inventoried plot mask
data(DTMParacou) # topography
data(CreekDistances) # relative elevation
data(SpeciesCriteria) # species exploitability criteria
data(ForestZoneVolumeParametersTable) # volume parameters
data(ParamCrownDiameterAllometry) # parameters values of the crown diameter allometry

library(LoggingLab)
```

# Simulation
```{r}
Rslt <- loggingsimulation(
  inventory = Paracou6_2016,
  plotmask = PlotMask, topography = DTMParacou,
  creekverticaldistance = CreekDistances$distvert,
  creekhorizontaldistance = CreekDistances$disthorz,
  speciescriteria = SpeciesCriteria,
  volumeparameters = ForestZoneVolumeParametersTable,
  scenario = "manual",
  objective = 30, fuel = "2", diversification = TRUE, winching = "2",
  directionalfelling = "2", specieslax = FALSE, objectivelax = TRUE,
  crowndiameterparameters = ParamCrownDiameterAllometry,
  advancedloggingparameters = loggingparameters(),
  iter = 30, cores = 15)

```

# Time
```{r}
SimulTime <- microbenchmark::microbenchmark(

  loggingsimulation(
    inventory = Paracou6_2016,
    plotmask = PlotMask, topography = DTMParacou,
    creekverticaldistance = CreekDistances$distvert,
    creekhorizontaldistance = CreekDistances$disthorz,
    speciescriteria = SpeciesCriteria,
    volumeparameters = ForestZoneVolumeParametersTable,
    scenario = "manual",
    objective = 30, fuel = "2", diversification = TRUE, winching = "2",
    directionalfelling = "2", specieslax = FALSE, objectivelax = TRUE,
    crowndiameterparameters = ParamCrownDiameterAllometry,
    advancedloggingparameters = loggingparameters(),
    iter = 30, cores = 15), times = 1L
)
```

# Summary
```{r}
loggingsummary(Rslt)
```

# Plot data prep
```{r}

NewInventory <- Rslt[[1]]$inventory
MainTrails <- Rslt[[1]]$MainTrails
HarvestablePolygons <- Rslt[[1]]$HarvestablePolygons
MachinePolygons <- Rslt[[1]]$MachinePolygons
SmoothedTrails <- Rslt[[1]]$SmoothedTrails
AdjustSmoothedTrails <- Rslt[[1]]$AdjustSmoothedTrails

NewInventory_crs <- NewInventory %>%
  getgeometry(TreePolygon) %>%
  sf::st_set_crs(sf::st_crs(MainTrails)) # set a crs

Harvestable <- sf::st_as_sf(
  dplyr::filter(NewInventory, LoggingStatus == "harvestable"),
  coords = c("Xutm", "Yutm")) %>%
  st_set_crs(st_crs(MainTrails))

HarvestableUp <- sf::st_as_sf(
  dplyr::filter(NewInventory, LoggingStatus == "harvestableUp"),
  coords = c("Xutm", "Yutm")) %>%
  st_set_crs(st_crs(MainTrails))

Selected <- sf::st_as_sf(
  dplyr::filter(NewInventory, Selected == "1"), coords = c("Xutm", "Yutm")) %>%
  st_set_crs(st_crs(MainTrails))

Reserve <- sf::st_as_sf(
  dplyr::filter(NewInventory, LoggingStatus == "reserve"),
  coords = c("Xutm", "Yutm")) %>%
  st_set_crs(st_crs(MainTrails))

Future <- sf::st_as_sf(
  dplyr::filter(NewInventory, LoggingStatus == "future"),
  coords = c("Xutm", "Yutm")) %>%
  st_set_crs(st_crs(MainTrails))

ProbedHollow <- sf::st_as_sf(
  dplyr::filter(NewInventory, ProbedHollow == "1"),
  coords = c("Xutm", "Yutm")) %>%
  st_set_crs(st_crs(MainTrails))

VisibleDefect <- sf::st_as_sf(
  dplyr::filter(NewInventory, VisibleDefect == "1"),
  coords = c("Xutm", "Yutm")) %>%
  st_set_crs(st_crs(MainTrails))

```

# Plot
```{r}
library(tidyverse)
library(sf)

ggplot() +

  labs(title = "Paracou P6 - RIL3 with fuelwood & hollow trees") +

  # MainTrails
  geom_sf(data = MainTrails,
          aes(color = "MainTrails")) +
  # Harvestable zones
  geom_sf(data = HarvestablePolygons,
          fill = "olivedrab", alpha = 0.1) +
  geom_sf(data = MachinePolygons,
          fill = "olivedrab", alpha = 0.5) +
  labs(alpha = "Harvestable") +

  geom_sf(data = VisibleDefect,
          aes(colour = "Visible defect"), show.legend = "point") +
  geom_sf(data = Future,
          aes(colour = "Future"), show.legend = "point", size = 4) +
  geom_sf(data = Reserve,
          aes(colour = "Reserve"), show.legend = "point", size = 4) +
  geom_sf(data = Harvestable,
          aes(colour = "Harvestable"), show.legend = "point", size = 4) +
  geom_sf(data = HarvestableUp,
          aes(colour = "HarvestableUp"), show.legend = "point", size = 4) +
  geom_sf(data = Selected,
          aes(colour = "Selected"), show.legend = "point") +
  geom_sf(data = ProbedHollow,
          aes(colour = "Probed hollow"), show.legend = "point") +

  # 2ndary trails
  geom_sf(data = st_as_sf(SmoothedTrails),
          aes(color = "Initial-trails"),alpha = 0.5) +
  # geom_sf(data = st_as_sf(SecondaryTrails$RawSecondTrails),
  #         color = "green",alpha = 0.5) +

  # 2ndary trails adjusted
  geom_sf(data = st_as_sf(AdjustSmoothedTrails),
          aes(color = "Adjusted-trails"),alpha = 0.5) +
  # geom_sf(data = st_as_sf(ScdTrailsAdj$RawSecondTrails),
  #         color = "red",alpha = 0.5) +

  # cuted trees
  geom_sf(data = NewInventory_crs, 
          alpha = 0.5, fill = "darkgreen") +

  scale_colour_manual(name = "Logging status",
                      values = c("Visible defect" = "pink",
                                 "Harvestable" = "skyblue",
                                 "HarvestableUp" = "blue", "Selected" = "red",
                                 "Future" = "orange",
                                 "Reserve" = "purple",
                                 "Probed hollow" = "forestgreen",
                                 "Harvestable area" = "olivedrab", 
                                 "Initial-trails" = "darkgreen" ,
                                 "Adjusted-trails" = "darkred",
                                 "MainTrails" = "red"))
# scale_fill_manual(name = "Logging structures",
#                   values = c("Harvestable area" = "olivedrab", 
#                              "Initial-trails" = "darkgreen" ,
#                              "Adjusted-trails" = "darkred",
#                              "MainTrails" = "red"))
```
```{r}
ggplot() +

  labs(title = "Paracou P6 - RIL3 with fuelwood & hollow trees") +

  # Harvestable zones
  geom_sf(data = HarvestablePolygons,
          fill = "olivedrab", alpha = 0.1) +
  geom_sf(data = MachinePolygons,
          fill = "olivedrab", alpha = 0.5) +
  labs(alpha = "Harvestable") +

  geom_sf(data = VisibleDefect,
          aes(colour = "Visible defect"), show.legend = "point") +
  geom_sf(data = Future,
          aes(colour = "Future"), show.legend = "point", size = 4) +
  geom_sf(data = Reserve,
          aes(colour = "Reserve"), show.legend = "point", size = 4) +
  geom_sf(data = Harvestable,
          aes(colour = "Harvestable"), show.legend = "point", size = 4) +
  geom_sf(data = HarvestableUp,
          aes(colour = "HarvestableUp"), show.legend = "point", size = 4) +
  geom_sf(data = Selected,
          aes(colour = "Selected"), show.legend = "point") +
  geom_sf(data = ProbedHollow,
          aes(colour = "Probed hollow"), show.legend = "point") +

  scale_colour_manual(name = "Logging status",
                      values = c("Visible defect" = "pink",
                                 "Harvestable" = "skyblue",
                                 "HarvestableUp" = "blue", "Selected" = "red",
                                 "Future" = "orange",
                                 "Reserve" = "purple",
                                 "Probed hollow" = "forestgreen"))
```

# Damage volume computing on Margaux's simulations
Plots (5): "SautLavilette", "Paracou11", "MontagnePlomb", "Acarouany", "Paracou16".  
Scenario: "2" = RIL2 without cable, "3" = RIL2 with cable  
Iterations number: 30 per plot.  
TotalResidualVolume: damage volume (m3) per harvestable hectare
```{r}
residual_volume_ha <- readr::read_csv("D:/VSC ManagFores/residual_volume_ha.csv")
residual_volume_ha
unique(residual_volume_ha$plot)
unique(residual_volume_ha$scenario)
```
## Per plot and scenario
```{r}
AverageResidualVolumes <- residual_volume_ha %>%
  group_by(plot, scenario) %>%
  mutate(AverageResidualVolume_ha = mean(TotalResidualVolume_ha)) %>%
  mutate(MaxResidualVolume_ha = max(TotalResidualVolume_ha)) %>%
  mutate(MinResidualVolume_ha = min(TotalResidualVolume_ha)) %>%
  select(plot,scenario,
         AverageResidualVolume_ha,
         MaxResidualVolume_ha, MinResidualVolume_ha) %>%
  unique()

AverageResidualVolumes
mean(AverageResidualVolumes$AverageResidualVolume_ha)
max(AverageResidualVolumes$AverageResidualVolume_ha)
min(AverageResidualVolumes$AverageResidualVolume_ha)
quantile(AverageResidualVolumes$AverageResidualVolume_ha)
```

## Per scenario
```{r}
AverageResidualVolumes_scenar <- residual_volume_ha %>%
  group_by(scenario) %>%
  mutate(AverageResidualVolume_ha = mean(TotalResidualVolume_ha)) %>%
  mutate(RV2.5 = quantile(TotalResidualVolume_ha, probs = 0.025)) %>%
  mutate(RV97.5 = quantile(TotalResidualVolume_ha, probs = 0.975)) %>%
  mutate(MaxResidualVolume_ha = max(TotalResidualVolume_ha)) %>%
  mutate(MinResidualVolume_ha = min(TotalResidualVolume_ha)) %>%
  select(scenario,
         AverageResidualVolume_ha,
         MaxResidualVolume_ha, MinResidualVolume_ha,
         RV2.5, RV97.5) %>%
  unique()

AverageResidualVolumes_scenar
mean(AverageResidualVolumes$AverageResidualVolume_ha)
quantile(AverageResidualVolumes$AverageResidualVolume_ha, probs = c(0.025, 0.975))
```